---
title: "BLMS Project: Airline Customer Satisfaction"
output: html_notebook
---

# Includes

```{r}
library(ggplot2)
library(tidyverse)
library(patchwork)
library(dplyr)
library(GGally)
```

# Initial Exploration of the Data

```{r}

table_raw <- read.csv("data.csv")

head(table_raw)

rows <- nrow(table_raw)
cols <- ncol(table_raw)

unique(table_raw$satisfaction)
unique(table_raw$Customer.Type)
unique(table_raw$Type.of.Travel)
unique(table_raw$Class)
# table(table_raw$Seat.comfort)
```

## Cleanup

Rename some of the columns / values to fix capitalization

```{r}
table <- table_raw

names(table)[names(table) == "satisfaction"                     ] <- "Satisfaction"
names(table)[names(table) == "Seat.comfort"                     ] <- "Seat.Comfort"
names(table)[names(table) == "Departure.Arrival.time.convenient"] <- "Departure.Arrival.Time.Convenient"
names(table)[names(table) == "Food.and.drink"                   ] <- "Food.and.Drink"
names(table)[names(table) == "Gate.location"                    ] <- "Gate.Location"
names(table)[names(table) == "Inflight.wifi.service"            ] <- "Inflight.WiFi"
names(table)[names(table) == "Inflight.entertainment"           ] <- "Inflight.Entertainment"
names(table)[names(table) == "Online.support"                   ] <- "Online.Support"
names(table)[names(table) == "Ease.of.Online.booking"           ] <- "Ease.of.Online.Booking"
names(table)[names(table) == "On.board.service"                 ] <- "On.Board.Service"
names(table)[names(table) == "Leg.room.service"                 ] <- "Leg.Room"
names(table)[names(table) == "Baggage.handling"                 ] <- "Baggage.Handling"
names(table)[names(table) == "Checkin.service"                  ] <- "Checkin"
names(table)[names(table) == "Online.boarding"                  ] <- "Online.Boarding"
names(table)[names(table) == "Departure.Delay.in.Minutes"       ] <- "Departure.Delay"
names(table)[names(table) == "Arrival.Delay.in.Minutes"         ] <- "Arrival.Delay"

table$Satisfaction  [table$Satisfaction   == "satisfied"        ] <- "Satisfied"
table$Satisfaction  [table$Satisfaction   == "dissatisfied"     ] <- "Dissatisfied"
table$Customer.Type [table$Customer.Type  == "Loyal Customer"   ] <- "Loyal"
table$Customer.Type [table$Customer.Type  == "disloyal Customer"] <- "Disloyal"
table$Type.of.Travel[table$Type.of.Travel == "Business travel"  ] <- "Businness"
table$Type.of.Travel[table$Type.of.Travel == "Personal Travel"  ] <- "Personal"

head(table)

```

| Name | Type | Values |
|-------------------------|------------------|-----------------------------|
| Satisfaction | categorical binary | 'Satisfied' / 'Dissatisfied' |
| Customer.Type | categorical binary | 'Loyal' / 'Disloyal' |
| Age | numeric discrete |  |
| Type.of.Travel | categorical binary | 'Business' / 'Personal’ |
| Class | categorical | 'Business' / 'Eco' / 'Eco Plus’ |
| Flight.Distance | numeric discrete | in kilometers |
| Seat.Comfort | score | from 0 to 5 |
| Departure.Arrival.Time.Convenient | score | from 0 to 5 |
| Food.and.Drink | score | from 0 to 5 |
| Gate.Location | score | from 0 to 5 |
| Inflight.WiFi | score | from 0 to 5 |
| Inflight.Entertainment | score | from 0 to 5 |
| Online.Support | score | from 0 to 5 |
| Ease.of.Online.Booking | score | from 0 to 5 |
| On.Board.Service | score | from 0 to 5 |
| Leg.Room | score | from 0 to 5 |
| Baggage.Handling | score | from 0 to 5 |
| Checkin | score | from 0 to 5 |
| Cleanliness | score | from 0 to 5 |
| Online.Boarding | score | from 0 to 5 |
| Departure.Delay | numeric discrete | in minutes |
| Arrival.Delay | numeric discrete | in minutes |

Inspect for NA

```{r}
table[!complete.cases(table), ]

# only 3 observations with NA in the arrival delay, just remove them from the dataset
table <- na.omit(table)
```

Define some useful constants

```{r}
score_vars <- c(
  "Seat.Comfort",
  "Departure.Arrival.Time.Convenient",
  "Food.and.Drink",
  "Gate.Location",
  "Inflight.WiFi",
  "Inflight.Entertainment",
  "Online.Support",
  "Ease.of.Online.Booking",
  "On.Board.Service",
  "Leg.Room",
  "Baggage.Handling",
  "Checkin",
  "Cleanliness",
  "Online.Boarding"
)
other_cat_vars <- c(
  "Customer.Type",
  "Type.of.Travel",
  "Class"
)
all_cat_vars <- c(other_cat_vars, score_vars)
num_vars <- c(
  "Age",
  "Flight.Distance",
  "Departure.Delay",
  "Arrival.Delay"
)
all_vars <- c(all_cat_vars, num_vars)
target_var <- "Satisfaction"
```

## Standardize

```{r}
table_std <- table %>%
  mutate(across(
    c(Departure.Delay, Arrival.Delay, Flight.Distance, Age),
    ~ as.numeric(scale(.))
  ))
```

## Visualization

### Boxplots

Boxplots of each variable against satisfaction. As expected satisfied customers show greater averages in the service quality scores and lower averages for delays.

```{r fig.width=14, fig.height=25}

pairs <- lapply(all_vars, function(v) {
  c("Satisfaction", v)
})

plots <- lapply(pairs, function(p) {
  cat_var <- p[1]
  num_var <- p[2]

  ggplot(table_std, aes(x = .data[[cat_var]], y = .data[[num_var]])) +
  geom_boxplot() +
  labs(
    x = cat_var,
    y = num_var,
    title =""
  ) +
  theme_bw()
})

wrap_plots(plots, ncol = 3)
```

### Barplots

Just to show the number of observations of each value of the categorical variables

```{r fig.width=14, fig.height=25}

plot_bar <- function(vars, scale = NULL) {
  x_scale <- if (!is.null(scale)) {
    scale_x_discrete(limits = scale)
  } else {
    NULL
  }
  
  plots <- lapply(vars, function(v) {
    ggplot(table, aes(
      x = factor(.data[[v]]),
      fill = factor(Satisfaction)
    )) +
      
    geom_bar() +
      
    x_scale +
    scale_fill_manual(
      values = c(
        "Dissatisfied" = "red",
        "Satisfied" = "forestgreen"
      )
    ) +
      
    labs(
      title = v,
      x = NULL,
      y = "Count",
      fill = "Satisfaction"
    ) +
    theme_bw()
  })
}


plots <- plot_bar(c(target_var, other_cat_vars))
plots <- c(plots, plot_bar(score_vars, c("0", "1", "2", "3", "4", "5")))

wrap_plots(plots, ncol = 3)
```

### Barplots Percentages

Here we show the percentages of satisfied/dissatisfied customers for each value of each categorical variable. This is somewhat more informative than the normal barplots, tho the trends can be approximated visually in the graphs above.

Note: there are some odd values, e.g. 100% of the customers that gave Leg.Room = 0 are satisfied, but notice that there is only 1 such observation, thus it does not represent a statistically relevant phenomena -\> almost all scores show few observations for score = 0, so generally ignore the corresponding column in the following graphs.

So excluding the first column, we see that the percentage of customers that gave score = x increases as x increases, as expected.

```{r fig.width=14, fig.height=20}

plot_bar_perc <- function(vars, scale = NULL) {
  x_scale <- if (!is.null(scale)) {
    scale_x_discrete(limits = scale)
  } else {
    NULL
  }
  
  plots <- lapply(vars, function(v) {

    ggplot(table, aes(
      x = factor(.data[[v]]),
      fill = Satisfaction
    )) +
    geom_bar(position = "fill") +
  
    geom_text(
      stat = "count",
      aes(label = after_stat(count)),
      position = position_fill(vjust = 0.5),
      color = "white",
      size = 3
    ) +
  
    scale_y_continuous(labels = scales::percent) +
    x_scale +
    scale_fill_manual(
      values = c(
        "Dissatisfied" = "red",
        "Satisfied" = "forestgreen"
      )
    ) +
      
    labs(
      title = v,
      x = NULL,
      y = "Percentage",
      fill = "Satisfaction"
    ) +
    theme_bw()
  })
}


plots <- c(plot_spacer(), plot_bar_perc(other_cat_vars))
plots <- c(plots, plot_bar_perc(score_vars, c("0", "1", "2", "3", "4", "5")))

wrap_plots(plots, ncol = 3)

```

### Pairplots

Using only the numerical variables, since the scores are just a whole lot of overlapping points.

Those are not very informative imo.

```{r fig.width=10, fig.height=10}

ggpairs(
  table_std,
  columns = num_vars,
  aes(color = Satisfaction, alpha = 0.5)
)
```
