---
title: "BLMS Project: Airline Customer Satisfaction-full dataset"
output: html_document
---


# Includes

```{r}
library(ggplot2)
library(tidyverse)
library(patchwork)
library(dplyr)
library(GGally)
library(corrplot)
library(BAS)
library(pROC)
```

### Initial cleanup

```{r}

table_raw <- read.csv("data_full.csv")


rows <- nrow(table_raw)
cols <- ncol(table_raw)

## Cleanup
# Rename some of the columns / values to fix capitalization

table <- table_raw

names(table)[names(table) == "satisfaction"                     ] <- "Satisfaction"
names(table)[names(table) == "Seat.comfort"                     ] <- "Seat.Comfort"
names(table)[names(table) == "Departure.Arrival.time.convenient"] <- "Departure.Arrival.Time.Convenient"
names(table)[names(table) == "Food.and.drink"                   ] <- "Food.and.Drink"
names(table)[names(table) == "Gate.location"                    ] <- "Gate.Location"
names(table)[names(table) == "Inflight.wifi.service"            ] <- "Inflight.WiFi"
names(table)[names(table) == "Inflight.entertainment"           ] <- "Inflight.Entertainment"
names(table)[names(table) == "Online.support"                   ] <- "Online.Support"
names(table)[names(table) == "Ease.of.Online.booking"           ] <- "Ease.of.Online.Booking"
names(table)[names(table) == "On.board.service"                 ] <- "On.Board.Service"
names(table)[names(table) == "Leg.room.service"                 ] <- "Leg.Room"
names(table)[names(table) == "Baggage.handling"                 ] <- "Baggage.Handling"
names(table)[names(table) == "Checkin.service"                  ] <- "Checkin"
names(table)[names(table) == "Online.boarding"                  ] <- "Online.Boarding"
names(table)[names(table) == "Departure.Delay.in.Minutes"       ] <- "Departure.Delay"
names(table)[names(table) == "Arrival.Delay.in.Minutes"         ] <- "Arrival.Delay"

table$Satisfaction  [table$Satisfaction   == "satisfied"        ] <- "Satisfied"
table$Satisfaction  [table$Satisfaction   == "dissatisfied"     ] <- "Dissatisfied"
table$Customer.Type [table$Customer.Type  == "Loyal Customer"   ] <- "Loyal"
table$Customer.Type [table$Customer.Type  == "disloyal Customer"] <- "Disloyal"
table$Type.of.Travel[table$Type.of.Travel == "Business travel"  ] <- "Businness"
table$Type.of.Travel[table$Type.of.Travel == "Personal Travel"  ] <- "Personal"
# Inspect for NA

table[!complete.cases(table), ]

# remove rows with NA (likely more than 3 here, since this is the full ~130k dataset
# rather than the cleaned 1000-row subset -- check the count before assuming it's negligible)
cat("Rows with NA:", sum(!complete.cases(table)), "out of", nrow(table), "\n")
table <- na.omit(table)
# Define some useful constants

score_vars <- c(
  "Seat.Comfort",
  "Departure.Arrival.Time.Convenient",
  "Food.and.Drink",
  "Gate.Location",
  "Inflight.WiFi",
  "Inflight.Entertainment",
  "Online.Support",
  "Ease.of.Online.Booking",
  "On.Board.Service",
  "Leg.Room",
  "Baggage.Handling",
  "Checkin",
  "Cleanliness",
  "Online.Boarding"
)
cat_vars <- c(
  "Customer.Type",
  "Type.of.Travel",
  "Class"
)
num_vars <- c(
  "Age",
  "Flight.Distance",
  "Departure.Delay",
  "Arrival.Delay"
)
all_vars <- c(score_vars, cat_vars, num_vars)
target_var <- "Satisfaction"
```

### Plots

```{r}
## Quick standardization for plotting

table_full_std <- table %>%
  mutate(across(
    c(Departure.Delay, Arrival.Delay, Flight.Distance, Age),
    ~ as.numeric(scale(.))
  ))
```

```{r fig.width=14, fig.height=20}
### Boxplots

# With ~130k rows, full boxplots per variable are still meaningful (boxplots
# summarize, they don't plot every point), so this scales fine.

pairs <- lapply(all_vars, function(v) {
  c("Satisfaction", v)
})

plots <- lapply(pairs, function(p) {
  cat_var <- p[1]
  num_var <- p[2]

  ggplot(table_full_std, aes(x = .data[[cat_var]], y = .data[[num_var]])) +
  geom_boxplot(outlier.alpha = 0.05) +
  labs(
    x = cat_var,
    y = num_var,
    title = ""
  ) +
  theme_bw()
})

wrap_plots(plots, ncol = 3)

```

```{r fig.width=14, fig.height=20}
### Barplots

# Counts of each value of the categorical/score variables, split by Satisfaction

plot_bar <- function(vars, scale = NULL) {
  x_scale <- if (!is.null(scale)) {
    scale_x_discrete(limits = scale)
  } else {
    NULL
  }
  
  plots <- lapply(vars, function(v) {
    ggplot(table, aes(
      x = factor(.data[[v]]),
      fill = factor(Satisfaction)
    )) +
      
    geom_bar() +
      
    x_scale +
    scale_fill_manual(
      values = c(
        "Dissatisfied" = "red",
        "Satisfied" = "forestgreen"
      )
    ) +
      
    labs(
      title = v,
      x = NULL,
      y = "Count",
      fill = "Satisfaction"
    ) +
    theme_bw()
  })
}

plots <- plot_bar(c(target_var, cat_vars))
plots <- c(plots, plot_bar(score_vars, c("0", "1", "2", "3", "4", "5")))

wrap_plots(plots, ncol = 3)

```

```{r fig.width=14, fig.height=20}
### Barplots Percentages

# Percentage of satisfied/dissatisfied for each value of each categorical/score
# variable -- more informative than raw counts since it normalizes class imbalance

plot_bar_perc <- function(vars, scale = NULL) {
  x_scale <- if (!is.null(scale)) {
    scale_x_discrete(limits = scale)
  } else {
    NULL
  }
  
  plots <- lapply(vars, function(v) {

    ggplot(table, aes(
      x = factor(.data[[v]]),
      fill = Satisfaction
    )) +
    geom_bar(position = "fill") +
  
    geom_text(
      stat = "count",
      aes(label = after_stat(count)),
      position = position_fill(vjust = 0.5),
      color = "white",
      size = 2.5
    ) +
  
    scale_y_continuous(labels = scales::percent) +
    x_scale +
    scale_fill_manual(
      values = c(
        "Dissatisfied" = "red",
        "Satisfied" = "forestgreen"
      )
    ) +
      
    labs(
      title = v,
      x = NULL,
      y = "Percentage",
      fill = "Satisfaction"
    ) +
    theme_bw()
  })
}

plots <- c(plot_spacer(), plot_bar_perc(cat_vars))
plots <- c(plots, plot_bar_perc(score_vars, c("0", "1", "2", "3", "4", "5")))

wrap_plots(plots, ncol = 3)
```

```{r fig.width=12, fig.height=10}

## Correlation

corr_matrix <- cor(table_full_std[, c(num_vars, score_vars)], use = "complete.obs")

corrplot(corr_matrix,
         method = "color",
         type = "upper",
         addCoef.col = "black",
         tl.col = "black",
         tl.srt = 45,
         number.cex = 0.7)

```

There are not many differences compared to the smaller version in the plots. The most notable ones is in the boxplot. Since we have more data, we are more likely to encounter data found at the tail of the distribution, this gives us more outlier, which are also further from the mean.

### Data preparation

```{r}
## Train-Test Split

# We use a random sample as the test set and the remaining rows as the learning
# set, fixing the seed so the split is reproducible.

# Perform train-test split at the beginning of the analysis since even variable
# selection based on a model trained on the whole data would be a data leakage
# when measuring performance on the test set.

set.seed(67)
idx <- sample(nrow(table))

n_test <- 39000 #~30 % of the data set

table_test  <- table[idx[1:n_test], ]
table_train <- table[idx[(n_test + 1):nrow(table)], ]

rownames(table_train) <- seq_len(nrow(table_train))
rownames(table_test)  <- seq_len(nrow(table_test))

cat("Learning set size:", nrow(table_train), "\n")
cat("Test set size:    ", nrow(table_test ), "\n")
## Standardize and Encode

# Compute mean and sd on TRAIN only
train_means <- sapply(table_train[, num_vars], mean)
train_sds   <- sapply(table_train[, num_vars], sd)

# Scale train
table_std <- table_train
table_std[, num_vars] <- scale(table_train[, num_vars], center = train_means, scale = train_sds)

# Scale test using the same (train-derived) parameters
test_std <- table_test
test_std[, num_vars] <- scale(table_test[, num_vars], center = train_means, scale = train_sds)

# Encode strings as factors -- BAS treats factors as categorical, introducing
# dummy variables for the n-1 possible values
table_enc <- table_std %>%
  mutate(across(where(is.character), as.factor))
test_enc  <- test_std %>%
  mutate(across(where(is.character), as.factor))
```

### The initial regression model

This takes a while to run     
It uses three different priors:      
bic prior -\> bayes information criterion, I believe we have seen that in AS, it is a large-sample approximation to the Bayes factor. AS I understand it, it is not a real prior, it just uses the BIC formula for the likelihood, it is also somehow connected to the g prior.     
aic prior -\> same idea, but uses AIC (again form AS??? Idk, I barely passed it)      
robust -\> proper prior, uses a mixture of the g priors, to be robust on the choice of g priors     
```{r}
fit_bic <- bas.glm(
  Satisfaction ~ Age + Seat.Comfort + Flight.Distance + Class + Departure.Delay + Arrival.Delay,
  data      = table_enc,
  family    = binomial(),
  betaprior = bic.prior(n = nrow(table_enc))
)

fit_aic <- bas.glm(
  Satisfaction ~ Age + Seat.Comfort + Flight.Distance + Class + Departure.Delay + Arrival.Delay,
  data      = table_enc,
  family    = binomial(),
  betaprior = aic.prior()
)

fit_robust <- bas.glm(
  Satisfaction ~ Age + Seat.Comfort + Flight.Distance + Class + Departure.Delay + Arrival.Delay,
  data      = table_enc,
  family    = binomial(),
  betaprior = robust()
)

pip_comparison <- data.frame(
  variable = fit_bic$namesx,
  PIP_BIC  = fit_bic$probne0,
  PIP_AIC  = fit_aic$probne0,
  PIP_robust = fit_robust$probne0
)
pip_comparison <- pip_comparison[order(pip_comparison$PIP_BIC, decreasing = TRUE), ]
pip_comparison

```

```{r}
# Pseudo-R² for a bas.glm fit
pseudo_r2 <- function(fit, data, prior_name) {
  
  # Fitted probabilities under BMA
  pred <- predict(fit, newdata = data, estimator = "BMA", type = "response")
  p_hat <- pred$fit
  
  y <- as.numeric(data$Satisfaction == "Satisfied")
  
  # McFadden R²
  # log-likelihood of fitted model
  ll_fit  <- sum(y * log(p_hat + 1e-10) + (1 - y) * log(1 - p_hat + 1e-10))
  # log-likelihood of null model (intercept only, predicts base rate everywhere)
  p_null  <- mean(y)
  ll_null <- sum(y * log(p_null) + (1 - y) * log(1 - p_null))
  mcfadden <- 1 - ll_fit / ll_null
  
  # Tjur R²
  mean_sat   <- mean(p_hat[y == 1])
  mean_disat <- mean(p_hat[y == 0])
  tjur <- mean_sat - mean_disat
  
  data.frame(
    prior    = prior_name,
    McFadden = round(mcfadden, 4),
    Tjur     = round(tjur,     4)
  )
}

r2_bic    <- pseudo_r2(fit_bic,    table_enc, "BIC")
r2_aic    <- pseudo_r2(fit_aic,    table_enc, "AIC")
r2_robust <- pseudo_r2(fit_robust, table_enc, "Robust")

r2_table <- rbind(r2_bic, r2_aic, r2_robust)
print(r2_table)
```

The R\^2 for the logistic regressions are hard to interpret on its own:    
<https://statisticalhorizons.com/r2logistic/> A funny article of guy arguing about which R\^2 is better.    
Ig we can just compare R\^2 between the models. Not as a value on its own.
```{r}
# Extract posterior means, SDs and credible intervals

extract_bas_summary <- function(fit, prior_name) {
  co <- coef(fit, estimator = "BMA")
  ci <- confint(co)
  
  data.frame(
    prior    = prior_name,
    variable = fit$namesx,
    PIP      = round(fit$probne0, 3),
    post_mean = round(co$postmean, 3),
    post_sd   = round(co$postsd,   3),
    lower_95  = round(ci[, 1],     3),
    upper_95  = round(ci[, 2],     3)
  )
}

summary_bic    <- extract_bas_summary(fit_bic,    "BIC")
summary_aic    <- extract_bas_summary(fit_aic,    "AIC")
summary_robust <- extract_bas_summary(fit_robust, "Robust")

# Stack all three
summary_all <- rbind(summary_bic, summary_aic, summary_robust)

# Wide comparison table: one row per variable, columns for each prior

make_wide_table <- function(summary_all, metric) {
  summary_all %>%
    select(prior, variable, all_of(metric)) %>%
    pivot_wider(names_from = prior, values_from = all_of(metric))
}

cat("=== Posterior Inclusion Probabilities ===\n")
print(make_wide_table(summary_all, "PIP"))

cat("\n=== Posterior Means ===\n")
print(make_wide_table(summary_all, "post_mean"))

cat("\n=== Posterior SDs ===\n")
print(make_wide_table(summary_all, "post_sd"))

cat("\n=== 95% Credible Interval Lower Bound ===\n")
print(make_wide_table(summary_all, "lower_95"))

cat("\n=== 95% Credible Interval Upper Bound ===\n")
print(make_wide_table(summary_all, "upper_95"))

```

```{r}
# Coefficient plot: posterior mean + 95% CI for each variable, 

plot_data <- summary_all %>%
  filter(variable != "Intercept")

ggplot(plot_data, aes(
  x    = post_mean,
  y    = reorder(variable, post_mean),
  xmin = lower_95,
  xmax = upper_95,
  color = prior
)) +
  geom_pointrange(
    position = position_dodge(width = 0.5),
    size     = 0.5
  ) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "red") +
  facet_wrap(~ prior, ncol = 3) +
  scale_color_manual(values = c(
    "BIC"    = "steelblue",
    "AIC"    = "darkorange",
    "Robust" = "forestgreen"
  )) +
  labs(
    title = "Posterior means and 95% credible intervals by prior",
    x     = "Posterior mean (BMA)",
    y     = NULL,
    color = "Prior"
  ) +
  theme_bw(base_size = 10) +
  theme(legend.position = "none")
```

```{r fig.width=12, fig.height=10}
# Posterior density plots for each coefficient under each prior

par(mfrow = c(3, 3))

plot(coef(fit_bic,    estimator = "BMA"), ask = FALSE)
mtext("BIC prior",    side = 3, line = -1.5, outer = TRUE, cex = 1.2)

par(mfrow = c(3, 3))

plot(coef(fit_aic,    estimator = "BMA"), ask = FALSE)
mtext("AIC prior",    side = 3, line = -1.5, outer = TRUE, cex = 1.2)

par(mfrow = c(3, 3))

plot(coef(fit_robust, estimator = "BMA"), ask = FALSE)
mtext("Robust prior", side = 3, line = -1.5, outer = TRUE, cex = 1.2)

par(mfrow = c(1, 1))
```

```{r}
compute_auc <- function(fit, train_data, test_data, prior_name) {
  
  y_train <- as.numeric(train_data$Satisfaction == "Satisfied")
  y_test  <- as.numeric(test_data$Satisfaction  == "Satisfied")
  
  # BMA predictions on train and test
  pred_train <- predict(fit, newdata = train_data, estimator = "BMA", type = "response")
  pred_test  <- predict(fit, newdata = test_data,  estimator = "BMA", type = "response")
  
  roc_train <- roc(y_train, pred_train$fit, quiet = TRUE)
  roc_test  <- roc(y_test,  pred_test$fit,  quiet = TRUE)
  
  data.frame(
    prior    = prior_name,
    AUC_train = round(as.numeric(auc(roc_train)), 4),
    AUC_test  = round(as.numeric(auc(roc_test)),  4)
  )
}

auc_bic    <- compute_auc(fit_bic,    table_enc, test_enc, "BIC")
auc_aic    <- compute_auc(fit_aic,    table_enc, test_enc, "AIC")
auc_robust <- compute_auc(fit_robust, table_enc, test_enc, "Robust")

auc_table <- rbind(auc_bic, auc_aic, auc_robust)
print(auc_table)
```
**How do they compare?**      
Robust and aic give more or less the same results, bic is somewhat different. This should be a results of the large data set. bic is the only one that takes into the account the size of the data set (as log(n)), which is significant here. It therefore is the best one for this situation.     
In terms of AUC, all the models perform the same.

**G prior**   
Another stability test could be with the g prior and different hyper parameters:
```{r}
fit_g1 <- bas.glm(
  Satisfaction ~ Age + Seat.Comfort + Flight.Distance + Class + Departure.Delay + Arrival.Delay,
  data      = table_enc,
  family    = binomial(),
  betaprior = g.prior(g = 1)
)

fit_g10 <- bas.glm(
  Satisfaction ~ Age + Seat.Comfort + Flight.Distance + Class + Departure.Delay + Arrival.Delay,
  data      = table_enc,
  family    = binomial(),
  betaprior = g.prior(g = 10)
)

fit_g100 <- bas.glm(
  Satisfaction ~ Age + Seat.Comfort + Flight.Distance + Class + Departure.Delay + Arrival.Delay,
  data      = table_enc,
  family    = binomial(),
  betaprior = g.prior(g = 100)
)

pip_comparison <- data.frame(
  variable = fit_g1$namesx,
  PIP_g1  = fit_g1$probne0,
  PIP_g10  = fit_g10$probne0,
  PIP_100 = fit_g100$probne0
)
pip_comparison <- pip_comparison[order(pip_comparison$PIP_g1, decreasing = TRUE), ]
pip_comparison

```

```{r}
extract_bas_summary <- function(fit, prior_name) {
  co <- coef(fit, estimator = "BMA")
  ci <- confint(co)
  
  data.frame(
    prior    = prior_name,
    variable = fit$namesx,
    PIP      = round(fit$probne0, 3),
    post_mean = round(co$postmean, 3),
    post_sd   = round(co$postsd,   3),
    lower_95  = round(ci[, 1],     3),
    upper_95  = round(ci[, 2],     3)
  )
}

summary_g1    <- extract_bas_summary(fit_g1,    "g1")
summary_g10    <- extract_bas_summary(fit_g10,    "g10")
summary_g100 <- extract_bas_summary(fit_g100, "g100")

# Stack all three
summary_all <- rbind(summary_g1, summary_g10, summary_g100)

# Wide comparison table: one row per variable, columns for each prior

make_wide_table <- function(summary_all, metric) {
  summary_all %>%
    select(prior, variable, all_of(metric)) %>%
    pivot_wider(names_from = prior, values_from = all_of(metric))
}

cat("=== Posterior Inclusion Probabilities ===\n")
print(make_wide_table(summary_all, "PIP"))

cat("\n=== Posterior Means ===\n")
print(make_wide_table(summary_all, "post_mean"))

cat("\n=== Posterior SDs ===\n")
print(make_wide_table(summary_all, "post_sd"))

cat("\n=== 95% Credible Interval Lower Bound ===\n")
print(make_wide_table(summary_all, "lower_95"))

cat("\n=== 95% Credible Interval Upper Bound ===\n")
print(make_wide_table(summary_all, "upper_95"))

```

```{r}
compute_auc <- function(fit, train_data, test_data, prior_name) {
  
  y_train <- as.numeric(train_data$Satisfaction == "Satisfied")
  y_test  <- as.numeric(test_data$Satisfaction  == "Satisfied")
  
  # BMA predictions on train and test
  pred_train <- predict(fit, newdata = train_data, estimator = "BMA", type = "response")
  pred_test  <- predict(fit, newdata = test_data,  estimator = "BMA", type = "response")
  
  roc_train <- roc(y_train, pred_train$fit, quiet = TRUE)
  roc_test  <- roc(y_test,  pred_test$fit,  quiet = TRUE)
  
  data.frame(
    prior    = prior_name,
    AUC_train = round(as.numeric(auc(roc_train)), 4),
    AUC_test  = round(as.numeric(auc(roc_test)),  4)
  )
}

auc_g1    <- compute_auc(fit_g1,    table_enc, test_enc, "g1")
auc_g10    <- compute_auc(fit_g10,    table_enc, test_enc, "g10")
auc_g100 <- compute_auc(fit_g100, table_enc, test_enc, "g100")

auc_table <- rbind(auc_g1, auc_g10, auc_g100)
print(auc_table)
```

model selection is robust to g (same variables are always included), but coefficient estimates are sensitive to g through the shrinkage mechanism, and g=100 or g=n is preferred here because with 130k observations there's no need for strong regularization, 'cause the data are informative enough on their own. So, 1 is very different, but 10 and 100 are more similar. AUC is again the same
### The PIP regression model
Run the same analysis but with variables chosen by PIP on the small dataset

```{r}
fit_bic <- bas.glm(
  Satisfaction ~
    Customer.Type + Inflight.Entertainment + Ease.of.Online.Booking + Class +
    Seat.Comfort + On.Board.Service + Type.of.Travel + Arrival.Delay + Food.and.Drink,
  data      = table_enc,
  family    = binomial(),
  betaprior = bic.prior(n = nrow(table_enc))
)

fit_aic <- bas.glm(
  Satisfaction ~
    Customer.Type + Inflight.Entertainment + Ease.of.Online.Booking + Class +
    Seat.Comfort + On.Board.Service + Type.of.Travel + Arrival.Delay + Food.and.Drink,
  data      = table_enc,
  family    = binomial(),
  betaprior = aic.prior()
)

fit_robust <- bas.glm(
  Satisfaction ~
    Customer.Type + Inflight.Entertainment + Ease.of.Online.Booking + Class +
    Seat.Comfort + On.Board.Service + Type.of.Travel + Arrival.Delay + Food.and.Drink,
  data      = table_enc,
  family    = binomial(),
  betaprior = robust()
)

pip_comparison <- data.frame(
  variable = fit_bic$namesx,
  PIP_BIC  = fit_bic$probne0,
  PIP_AIC  = fit_aic$probne0,
  PIP_robust = fit_robust$probne0
)
pip_comparison <- pip_comparison[order(pip_comparison$PIP_BIC, decreasing = TRUE), ]
pip_comparison

```

```{r}
# Pseudo-R² for a bas.glm fit
pseudo_r2 <- function(fit, data, prior_name) {
  
  # Fitted probabilities under BMA
  pred <- predict(fit, newdata = data, estimator = "BMA", type = "response")
  p_hat <- pred$fit
  
  y <- as.numeric(data$Satisfaction == "Satisfied")
  
  # McFadden R²
  # log-likelihood of fitted model
  ll_fit  <- sum(y * log(p_hat + 1e-10) + (1 - y) * log(1 - p_hat + 1e-10))
  # log-likelihood of null model (intercept only, predicts base rate everywhere)
  p_null  <- mean(y)
  ll_null <- sum(y * log(p_null) + (1 - y) * log(1 - p_null))
  mcfadden <- 1 - ll_fit / ll_null
  
  # Tjur R²
  mean_sat   <- mean(p_hat[y == 1])
  mean_disat <- mean(p_hat[y == 0])
  tjur <- mean_sat - mean_disat
  
  data.frame(
    prior    = prior_name,
    McFadden = round(mcfadden, 4),
    Tjur     = round(tjur,     4)
  )
}

r2_bic    <- pseudo_r2(fit_bic,    table_enc, "BIC")
r2_aic    <- pseudo_r2(fit_aic,    table_enc, "AIC")
r2_robust <- pseudo_r2(fit_robust, table_enc, "Robust")

r2_table <- rbind(r2_bic, r2_aic, r2_robust)
print(r2_table)
```

```{r}
# Extract posterior means, SDs and credible intervals

extract_bas_summary <- function(fit, prior_name) {
  co <- coef(fit, estimator = "BMA")
  ci <- confint(co)
  
  data.frame(
    prior    = prior_name,
    variable = fit$namesx,
    PIP      = round(fit$probne0, 3),
    post_mean = round(co$postmean, 3),
    post_sd   = round(co$postsd,   3),
    lower_95  = round(ci[, 1],     3),
    upper_95  = round(ci[, 2],     3)
  )
}

summary_bic    <- extract_bas_summary(fit_bic,    "BIC")
summary_aic    <- extract_bas_summary(fit_aic,    "AIC")
summary_robust <- extract_bas_summary(fit_robust, "Robust")

# Stack all three
summary_all <- rbind(summary_bic, summary_aic, summary_robust)

# Wide comparison table: one row per variable, columns for each prior

make_wide_table <- function(summary_all, metric) {
  summary_all %>%
    select(prior, variable, all_of(metric)) %>%
    pivot_wider(names_from = prior, values_from = all_of(metric))
}

cat("=== Posterior Inclusion Probabilities ===\n")
print(make_wide_table(summary_all, "PIP"))

cat("\n=== Posterior Means ===\n")
print(make_wide_table(summary_all, "post_mean"))

cat("\n=== Posterior SDs ===\n")
print(make_wide_table(summary_all, "post_sd"))

cat("\n=== 95% Credible Interval Lower Bound ===\n")
print(make_wide_table(summary_all, "lower_95"))

cat("\n=== 95% Credible Interval Upper Bound ===\n")
print(make_wide_table(summary_all, "upper_95"))

```

```{r}
# Coefficient plot: posterior mean + 95% CI for each variable, 

plot_data <- summary_all %>%
  filter(variable != "Intercept")

ggplot(plot_data, aes(
  x    = post_mean,
  y    = reorder(variable, post_mean),
  xmin = lower_95,
  xmax = upper_95,
  color = prior
)) +
  geom_pointrange(
    position = position_dodge(width = 0.5),
    size     = 0.5
  ) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "red") +
  facet_wrap(~ prior, ncol = 3) +
  scale_color_manual(values = c(
    "BIC"    = "steelblue",
    "AIC"    = "darkorange",
    "Robust" = "forestgreen"
  )) +
  labs(
    title = "Posterior means and 95% credible intervals by prior",
    x     = "Posterior mean (BMA)",
    y     = NULL,
    color = "Prior"
  ) +
  theme_bw(base_size = 10) +
  theme(legend.position = "none")
```

```{r fig.width=12, fig.height=10}
# Posterior density plots for each coefficient under each prior

par(mfrow = c(3, 3))

plot(coef(fit_bic,    estimator = "BMA"), ask = FALSE)
mtext("BIC prior",    side = 3, line = -1.5, outer = TRUE, cex = 1.2)

par(mfrow = c(3, 3))

plot(coef(fit_aic,    estimator = "BMA"), ask = FALSE)
mtext("AIC prior",    side = 3, line = -1.5, outer = TRUE, cex = 1.2)

par(mfrow = c(3, 3))

plot(coef(fit_robust, estimator = "BMA"), ask = FALSE)
mtext("Robust prior", side = 3, line = -1.5, outer = TRUE, cex = 1.2)

par(mfrow = c(1, 1))
```

```{r}
compute_auc <- function(fit, train_data, test_data, prior_name) {
  
  y_train <- as.numeric(train_data$Satisfaction == "Satisfied")
  y_test  <- as.numeric(test_data$Satisfaction  == "Satisfied")
  
  # BMA predictions on train and test
  pred_train <- predict(fit, newdata = train_data, estimator = "BMA", type = "response")
  pred_test  <- predict(fit, newdata = test_data,  estimator = "BMA", type = "response")
  
  roc_train <- roc(y_train, pred_train$fit, quiet = TRUE)
  roc_test  <- roc(y_test,  pred_test$fit,  quiet = TRUE)
  
  data.frame(
    prior    = prior_name,
    AUC_train = round(as.numeric(auc(roc_train)), 4),
    AUC_test  = round(as.numeric(auc(roc_test)),  4)
  )
}

auc_bic    <- compute_auc(fit_bic,    table_enc, test_enc, "BIC")
auc_aic    <- compute_auc(fit_aic,    table_enc, test_enc, "AIC")
auc_robust <- compute_auc(fit_robust, table_enc, test_enc, "Robust")

auc_table <- rbind(auc_bic, auc_aic, auc_robust)
print(auc_table)
```

**G prior**   
```{r}
fit_g1 <- bas.glm(
  Satisfaction ~
    Customer.Type + Inflight.Entertainment + Ease.of.Online.Booking + Class +
    Seat.Comfort + On.Board.Service + Type.of.Travel + Arrival.Delay + Food.and.Drink,
  data      = table_enc,
  family    = binomial(),
  betaprior = g.prior(g = 1)
)

fit_g10 <- bas.glm(
  Satisfaction ~
    Customer.Type + Inflight.Entertainment + Ease.of.Online.Booking + Class +
    Seat.Comfort + On.Board.Service + Type.of.Travel + Arrival.Delay + Food.and.Drink,
  data      = table_enc,
  family    = binomial(),
  betaprior = g.prior(g = 10)
)

fit_g100 <- bas.glm(
  Satisfaction ~
    Customer.Type + Inflight.Entertainment + Ease.of.Online.Booking + Class +
    Seat.Comfort + On.Board.Service + Type.of.Travel + Arrival.Delay + Food.and.Drink,
  data      = table_enc,
  family    = binomial(),
  betaprior = g.prior(g = 100)
)

pip_comparison <- data.frame(
  variable = fit_g1$namesx,
  PIP_g1  = fit_g1$probne0,
  PIP_g10  = fit_g10$probne0,
  PIP_100 = fit_g100$probne0
)
pip_comparison <- pip_comparison[order(pip_comparison$PIP_g1, decreasing = TRUE), ]
pip_comparison

```

```{r}
extract_bas_summary <- function(fit, prior_name) {
  co <- coef(fit, estimator = "BMA")
  ci <- confint(co)
  
  data.frame(
    prior    = prior_name,
    variable = fit$namesx,
    PIP      = round(fit$probne0, 3),
    post_mean = round(co$postmean, 3),
    post_sd   = round(co$postsd,   3),
    lower_95  = round(ci[, 1],     3),
    upper_95  = round(ci[, 2],     3)
  )
}

summary_g1    <- extract_bas_summary(fit_g1,    "g1")
summary_g10    <- extract_bas_summary(fit_g10,    "g10")
summary_g100 <- extract_bas_summary(fit_g100, "g100")

# Stack all three
summary_all <- rbind(summary_g1, summary_g10, summary_g100)

# Wide comparison table: one row per variable, columns for each prior

make_wide_table <- function(summary_all, metric) {
  summary_all %>%
    select(prior, variable, all_of(metric)) %>%
    pivot_wider(names_from = prior, values_from = all_of(metric))
}

cat("=== Posterior Inclusion Probabilities ===\n")
print(make_wide_table(summary_all, "PIP"))

cat("\n=== Posterior Means ===\n")
print(make_wide_table(summary_all, "post_mean"))

cat("\n=== Posterior SDs ===\n")
print(make_wide_table(summary_all, "post_sd"))

cat("\n=== 95% Credible Interval Lower Bound ===\n")
print(make_wide_table(summary_all, "lower_95"))

cat("\n=== 95% Credible Interval Upper Bound ===\n")
print(make_wide_table(summary_all, "upper_95"))

```

```{r}
compute_auc <- function(fit, train_data, test_data, prior_name) {
  
  y_train <- as.numeric(train_data$Satisfaction == "Satisfied")
  y_test  <- as.numeric(test_data$Satisfaction  == "Satisfied")
  
  # BMA predictions on train and test
  pred_train <- predict(fit, newdata = train_data, estimator = "BMA", type = "response")
  pred_test  <- predict(fit, newdata = test_data,  estimator = "BMA", type = "response")
  
  roc_train <- roc(y_train, pred_train$fit, quiet = TRUE)
  roc_test  <- roc(y_test,  pred_test$fit,  quiet = TRUE)
  
  data.frame(
    prior    = prior_name,
    AUC_train = round(as.numeric(auc(roc_train)), 4),
    AUC_test  = round(as.numeric(auc(roc_test)),  4)
  )
}

auc_g1    <- compute_auc(fit_g1,    table_enc, test_enc, "g1")
auc_g10    <- compute_auc(fit_g10,    table_enc, test_enc, "g10")
auc_g100 <- compute_auc(fit_g100, table_enc, test_enc, "g100")

auc_table <- rbind(auc_g1, auc_g10, auc_g100)
print(auc_table)
```


