---
title: "BLMS Project: Airline Customer Satisfaction"
output: html_notebook
---

# Includes

```{r}
library(ggplot2)
library(tidyverse)
library(patchwork)
library(dplyr)
library(GGally)
library(corrplot)
library(BAS)
```

# Initial Exploration of the Data

```{r}

table_raw <- read.csv("data.csv")

head(table_raw)

rows <- nrow(table_raw)
cols <- ncol(table_raw)

unique(table_raw$satisfaction)
unique(table_raw$Customer.Type)
unique(table_raw$Type.of.Travel)
unique(table_raw$Class)
```

## Cleanup

Rename some of the columns / values to fix capitalization

```{r}
table <- table_raw

names(table)[names(table) == "satisfaction"                     ] <- "Satisfaction"
names(table)[names(table) == "Seat.comfort"                     ] <- "Seat.Comfort"
names(table)[names(table) == "Departure.Arrival.time.convenient"] <- "Departure.Arrival.Time.Convenient"
names(table)[names(table) == "Food.and.drink"                   ] <- "Food.and.Drink"
names(table)[names(table) == "Gate.location"                    ] <- "Gate.Location"
names(table)[names(table) == "Inflight.wifi.service"            ] <- "Inflight.WiFi"
names(table)[names(table) == "Inflight.entertainment"           ] <- "Inflight.Entertainment"
names(table)[names(table) == "Online.support"                   ] <- "Online.Support"
names(table)[names(table) == "Ease.of.Online.booking"           ] <- "Ease.of.Online.Booking"
names(table)[names(table) == "On.board.service"                 ] <- "On.Board.Service"
names(table)[names(table) == "Leg.room.service"                 ] <- "Leg.Room"
names(table)[names(table) == "Baggage.handling"                 ] <- "Baggage.Handling"
names(table)[names(table) == "Checkin.service"                  ] <- "Checkin"
names(table)[names(table) == "Online.boarding"                  ] <- "Online.Boarding"
names(table)[names(table) == "Departure.Delay.in.Minutes"       ] <- "Departure.Delay"
names(table)[names(table) == "Arrival.Delay.in.Minutes"         ] <- "Arrival.Delay"

table$Satisfaction  [table$Satisfaction   == "satisfied"        ] <- "Satisfied"
table$Satisfaction  [table$Satisfaction   == "dissatisfied"     ] <- "Dissatisfied"
table$Customer.Type [table$Customer.Type  == "Loyal Customer"   ] <- "Loyal"
table$Customer.Type [table$Customer.Type  == "disloyal Customer"] <- "Disloyal"
table$Type.of.Travel[table$Type.of.Travel == "Business travel"  ] <- "Businness"
table$Type.of.Travel[table$Type.of.Travel == "Personal Travel"  ] <- "Personal"

head(table)

```

| Name | Type | Values |
|----|----|----|
| Satisfaction | categorical binary | 'Satisfied' / 'Dissatisfied' |
| Customer.Type | categorical binary | 'Loyal' / 'Disloyal' |
| Age | numeric discrete |  |
| Type.of.Travel | categorical binary | 'Business' / 'Personal’ |
| Class | categorical | 'Business' / 'Eco' / 'Eco Plus’ |
| Flight.Distance | numeric discrete | in kilometers |
| Seat.Comfort | score | from 0 to 5 |
| Departure.Arrival.Time.Convenient | score | from 0 to 5 |
| Food.and.Drink | score | from 0 to 5 |
| Gate.Location | score | from 0 to 5 |
| Inflight.WiFi | score | from 0 to 5 |
| Inflight.Entertainment | score | from 0 to 5 |
| Online.Support | score | from 0 to 5 |
| Ease.of.Online.Booking | score | from 0 to 5 |
| On.Board.Service | score | from 0 to 5 |
| Leg.Room | score | from 0 to 5 |
| Baggage.Handling | score | from 0 to 5 |
| Checkin | score | from 0 to 5 |
| Cleanliness | score | from 0 to 5 |
| Online.Boarding | score | from 0 to 5 |
| Departure.Delay | numeric discrete | in minutes |
| Arrival.Delay | numeric discrete | in minutes |

Inspect for NA

```{r}
table[!complete.cases(table), ]

# only 3 observations with NA in the arrival delay, just remove them from the dataset
table <- na.omit(table)
```

Define some useful constants

```{r}
score_vars <- c(
  "Seat.Comfort",
  "Departure.Arrival.Time.Convenient",
  "Food.and.Drink",
  "Gate.Location",
  "Inflight.WiFi",
  "Inflight.Entertainment",
  "Online.Support",
  "Ease.of.Online.Booking",
  "On.Board.Service",
  "Leg.Room",
  "Baggage.Handling",
  "Checkin",
  "Cleanliness",
  "Online.Boarding"
)
other_cat_vars <- c(
  "Customer.Type",
  "Type.of.Travel",
  "Class"
)
all_cat_vars <- c(other_cat_vars, score_vars)
num_vars <- c(
  "Age",
  "Flight.Distance",
  "Departure.Delay",
  "Arrival.Delay"
)
all_vars <- c(all_cat_vars, num_vars)
target_var <- "Satisfaction"
```

## Standardize and Encode

```{r}
table_std <- table %>%
  mutate(across(
    c(Departure.Delay, Arrival.Delay, Flight.Distance, Age),
    ~ as.numeric(scale(.))
  ))

# the encoding in R can be done just by converting the column of strings to a column of Factor (which just maps the strings to ints)
# BAS will treat Factors as categorical variables, introducing the dummy variables for the n-1 possible values it can assume.
table_enc <- table_std %>%
  mutate(across(where(is.character), as.factor))
```

## Visualization

### Boxplots

Boxplots of each variable against satisfaction. As expected satisfied customers show greater averages in the service quality scores and lower averages for delays.

```{r fig.width=14, fig.height=25}

pairs <- lapply(all_vars, function(v) {
  c("Satisfaction", v)
})

plots <- lapply(pairs, function(p) {
  cat_var <- p[1]
  num_var <- p[2]

  ggplot(table_std, aes(x = .data[[cat_var]], y = .data[[num_var]])) +
  geom_boxplot() +
  labs(
    x = cat_var,
    y = num_var,
    title =""
  ) +
  theme_bw()
})

wrap_plots(plots, ncol = 3)
```

### Barplots

Just to show the number of observations of each value of the categorical variables

```{r fig.width=14, fig.height=25}

plot_bar <- function(vars, scale = NULL) {
  x_scale <- if (!is.null(scale)) {
    scale_x_discrete(limits = scale)
  } else {
    NULL
  }
  
  plots <- lapply(vars, function(v) {
    ggplot(table, aes(
      x = factor(.data[[v]]),
      fill = factor(Satisfaction)
    )) +
      
    geom_bar() +
      
    x_scale +
    scale_fill_manual(
      values = c(
        "Dissatisfied" = "red",
        "Satisfied" = "forestgreen"
      )
    ) +
      
    labs(
      title = v,
      x = NULL,
      y = "Count",
      fill = "Satisfaction"
    ) +
    theme_bw()
  })
}


plots <- plot_bar(c(target_var, other_cat_vars))
plots <- c(plots, plot_bar(score_vars, c("0", "1", "2", "3", "4", "5")))

wrap_plots(plots, ncol = 3)
```

### Barplots Percentages

Here we show the percentages of satisfied/dissatisfied customers for each value of each categorical variable. This is somewhat more informative than the normal barplots, tho the trends can be approximated visually in the graphs above.

Note: there are some odd values, e.g. 100% of the customers that gave Leg.Room = 0 are satisfied, but notice that there is only 1 such observation, thus it does not represent a statistically relevant phenomena -\> almost all scores show few observations for score = 0, so generally ignore the corresponding column in the following graphs.

So excluding the first column, we see that the percentage of customers that gave score = x increases as x increases, as expected.

```{r fig.width=14, fig.height=20}

plot_bar_perc <- function(vars, scale = NULL) {
  x_scale <- if (!is.null(scale)) {
    scale_x_discrete(limits = scale)
  } else {
    NULL
  }
  
  plots <- lapply(vars, function(v) {

    ggplot(table, aes(
      x = factor(.data[[v]]),
      fill = Satisfaction
    )) +
    geom_bar(position = "fill") +
  
    geom_text(
      stat = "count",
      aes(label = after_stat(count)),
      position = position_fill(vjust = 0.5),
      color = "white",
      size = 3
    ) +
  
    scale_y_continuous(labels = scales::percent) +
    x_scale +
    scale_fill_manual(
      values = c(
        "Dissatisfied" = "red",
        "Satisfied" = "forestgreen"
      )
    ) +
      
    labs(
      title = v,
      x = NULL,
      y = "Percentage",
      fill = "Satisfaction"
    ) +
    theme_bw()
  })
}


plots <- c(plot_spacer(), plot_bar_perc(other_cat_vars))
plots <- c(plots, plot_bar_perc(score_vars, c("0", "1", "2", "3", "4", "5")))

wrap_plots(plots, ncol = 3)

```

### Pairplots

Using only the numerical variables, since the scores are just a whole lot of overlapping points.

Those are not very informative imo.

```{r fig.width=10, fig.height=10}

ggpairs(
  table_std,
  columns = num_vars,
  aes(color = Satisfaction, alpha = 0.5)
)
```

## Correlation

From the correlation matrix we see that most predictors are uncorrelated, with a few exceptions, such as

- between the "Online" sth sth variables, which kinda makes sense, a good online service is good across all aspects.

- between the flight enjoyment variables (Seat.Comfort, Food.and.Drink, Inflight.Entertainment)

and more with similar interpretations.

```{r fig.width=12, fig.height=12}

corr_matrix <- cor(table_std[, c(num_vars, score_vars)], use = "complete.obs")

corrplot(corr_matrix,
         method = "color",
         type = "upper",
         addCoef.col = "black",
         tl.col = "black",
         tl.srt = 45)
```

# Initial Regression Model

Start by fitting a regression model using only the predictors

$$\texttt{Age},\ \texttt{SeatComfort},\ \texttt{FlightDistance},\ \texttt{Class},\ \texttt{DepartureDelay},\ \texttt{ArrivalDelay}$$

First of all, BAS is using BMA (Bayesian Model Averaging), so we train many models and give them a weight.

To interpret the summary:

- P(B != 0 \| Y): aka Posterior Inclusion Probability (PIP), how likely the variable belongs to the true model

- The 1/0 under the models for the first rows (those corresponding to the predictors): boolean flag, whether that predictor is included in the model

- BF: Bayes Factor, idk what it is but it's a metric that compares each model with the best one, the higher the better, the best model has BF = 1 by construction.

- PostProb: probability that that model is the true one, given the data

- logmar: the higher the better

From the summary we see that:

- Flight.Distance and Departure.Delay are not selected very frequently across all the models trained by BAS, they might not capture useful data to fit the Satisfaction

- The best model obtained is much better than the others: PostProbs and BF are much higher than the other models, meaning that the data strongly prefers model 1 over the other alternatives.

- R2 is very bad for all the models, we are not fitting the data at all

In the summary of coefficients reports the average over the evaluated models weighted by their posterior probability, where

- post mean: average effect of a variable across all models, but models where the variable is absent count as 0 effect, so this is a shrunken, uncertainty-aware estimate.

- post SD: uncertainty on the effect, including the model uncertainty

so we can say that

- Age has a positive effect and high PIP: older passengers have higher probability of satisfaction; the same goes for Seat.Comfort, as expected

- Arrival.Delay has a negative effect (as expected) and discrete PIP, so there is some uncertainty in whether this predictor is actually important.

- Departure.Delay has almost 0 mean and very poor PIP, this is probably due to the fact that once Arrival.Delay is included, Departure.Delay adds little information.

- Flight.Distance: same as for Departure.Delay, this is probably irrelevant because people know the flight length in advance.

The graph of the residuals shows two parallel lines (the two target classes), but the red line isn't horizontal around the y = 0 line, so this model is garbage.

Note: all the values are generated with some randomness, so they can change slightly, but the interpretation should remain the same

```{r fig.width=8, fig.height=8}

fit <- bas.glm(
  Satisfaction ~ Age + Flight.Distance + Seat.Comfort + Departure.Delay + Arrival.Delay,
  data = table_enc, family = binomial()
)

summary(fit)
coef(fit)
print("--------------------------------------------", quote = FALSE)
confint(coef(fit))

plot(fit, which = 1)
plot(confint(coef(fit)),main="coeff confidence intervals")
```
