---
title: "BLMS Project: Airline Customer Satisfaction"
output: html_notebook
---

# Includes

```{r}
library(ggplot2)
library(tidyverse)
library(patchwork)
library(dplyr)
library(GGally)
library(corrplot)
library(BAS)
library(pROC)
```

# Initial Exploration of the Data

```{r}

table_raw <- read.csv("data.csv")

head(table_raw)

rows <- nrow(table_raw)
cols <- ncol(table_raw)

unique(table_raw$satisfaction)
unique(table_raw$Customer.Type)
unique(table_raw$Type.of.Travel)
unique(table_raw$Class)
```

## Cleanup

Rename some of the columns / values to fix capitalization

```{r}
table <- table_raw

names(table)[names(table) == "satisfaction"                     ] <- "Satisfaction"
names(table)[names(table) == "Seat.comfort"                     ] <- "Seat.Comfort"
names(table)[names(table) == "Departure.Arrival.time.convenient"] <- "Departure.Arrival.Time.Convenient"
names(table)[names(table) == "Food.and.drink"                   ] <- "Food.and.Drink"
names(table)[names(table) == "Gate.location"                    ] <- "Gate.Location"
names(table)[names(table) == "Inflight.wifi.service"            ] <- "Inflight.WiFi"
names(table)[names(table) == "Inflight.entertainment"           ] <- "Inflight.Entertainment"
names(table)[names(table) == "Online.support"                   ] <- "Online.Support"
names(table)[names(table) == "Ease.of.Online.booking"           ] <- "Ease.of.Online.Booking"
names(table)[names(table) == "On.board.service"                 ] <- "On.Board.Service"
names(table)[names(table) == "Leg.room.service"                 ] <- "Leg.Room"
names(table)[names(table) == "Baggage.handling"                 ] <- "Baggage.Handling"
names(table)[names(table) == "Checkin.service"                  ] <- "Checkin"
names(table)[names(table) == "Online.boarding"                  ] <- "Online.Boarding"
names(table)[names(table) == "Departure.Delay.in.Minutes"       ] <- "Departure.Delay"
names(table)[names(table) == "Arrival.Delay.in.Minutes"         ] <- "Arrival.Delay"

table$Satisfaction  [table$Satisfaction   == "satisfied"        ] <- "Satisfied"
table$Satisfaction  [table$Satisfaction   == "dissatisfied"     ] <- "Dissatisfied"
table$Customer.Type [table$Customer.Type  == "Loyal Customer"   ] <- "Loyal"
table$Customer.Type [table$Customer.Type  == "disloyal Customer"] <- "Disloyal"
table$Type.of.Travel[table$Type.of.Travel == "Business travel"  ] <- "Businness"
table$Type.of.Travel[table$Type.of.Travel == "Personal Travel"  ] <- "Personal"

head(table)

```

| Name | Type | Values |
|----|----|----|
| Satisfaction | categorical binary | 'Satisfied' / 'Dissatisfied' |
| Customer.Type | categorical binary | 'Loyal' / 'Disloyal' |
| Age | numeric discrete |  |
| Type.of.Travel | categorical binary | 'Business' / 'Personal’ |
| Class | categorical | 'Business' / 'Eco' / 'Eco Plus’ |
| Flight.Distance | numeric discrete | in kilometers |
| Seat.Comfort | score | from 0 to 5 |
| Departure.Arrival.Time.Convenient | score | from 0 to 5 |
| Food.and.Drink | score | from 0 to 5 |
| Gate.Location | score | from 0 to 5 |
| Inflight.WiFi | score | from 0 to 5 |
| Inflight.Entertainment | score | from 0 to 5 |
| Online.Support | score | from 0 to 5 |
| Ease.of.Online.Booking | score | from 0 to 5 |
| On.Board.Service | score | from 0 to 5 |
| Leg.Room | score | from 0 to 5 |
| Baggage.Handling | score | from 0 to 5 |
| Checkin | score | from 0 to 5 |
| Cleanliness | score | from 0 to 5 |
| Online.Boarding | score | from 0 to 5 |
| Departure.Delay | numeric discrete | in minutes |
| Arrival.Delay | numeric discrete | in minutes |

Inspect for NA

```{r}
table[!complete.cases(table), ]

# only 3 observations with NA in the arrival delay, just remove them from the dataset
table <- na.omit(table)
```

Define some useful constants

```{r}
score_vars <- c(
  "Seat.Comfort",
  "Departure.Arrival.Time.Convenient",
  "Food.and.Drink",
  "Gate.Location",
  "Inflight.WiFi",
  "Inflight.Entertainment",
  "Online.Support",
  "Ease.of.Online.Booking",
  "On.Board.Service",
  "Leg.Room",
  "Baggage.Handling",
  "Checkin",
  "Cleanliness",
  "Online.Boarding"
)
cat_vars <- c(
  "Customer.Type",
  "Type.of.Travel",
  "Class"
)
num_vars <- c(
  "Age",
  "Flight.Distance",
  "Departure.Delay",
  "Arrival.Delay"
)
all_vars <- c(score_vars, cat_vars, num_vars)
target_var <- "Satisfaction"
```

## Train-Test Split

We use the first `n_test` observations as the test set and the remaining ones as the learning set. This way the split is fully reproducible without relying on random sampling.

Perform train-test split at the beginning of the analysis since even variable selection based on a model trained on the whole data is then a data leakage when measuring the performance on the test set.

```{r}

# to make it random
set.seed(42)
idx <- sample(nrow(table))

table_test  <- table[idx[1:300], ]
table_train <- table[idx[301:nrow(table)], ]

#n_test    <- 400
#n_total   <- nrow(table_enc)
#n_learn   <- n_total - n_test

rownames(table_train) <- seq_len(nrow(table_train))
rownames(table_test)  <- seq_len(nrow(table_test))

cat("Learning set size:", nrow(table_train), "\n")
cat("Test set size:    ", nrow(table_test ), "\n")
```

## Standardize and Encode

```{r}

# Compute mean and sd on TRAIN only
train_means <- sapply(table_train[, num_vars], mean)
train_sds   <- sapply(table_train[, num_vars], sd)

# Scale train
table_std <- table_train
table_std[, num_vars] <- scale(table_train[, num_vars], center = train_means, scale = train_sds)

# Scale test using 🚂 parameters
test_std <- table_test
test_std[, num_vars] <- scale(table_test[, num_vars], center = train_means, scale = train_sds)

# the encoding in R can be done just by converting the column of strings to a column of Factor (which just maps the strings to ints)
# BAS will treat Factors as categorical variables, introducing the dummy variables for the n-1 possible values it can assume.
table_enc <- table_std %>%
  mutate(across(where(is.character), as.factor))
test_enc  <- test_std %>%
  mutate(across(where(is.character), as.factor))

# For other purposes we might still want to use the full dataset, like visualization
table_full_std <- table %>%
  mutate(across(
    c(Departure.Delay, Arrival.Delay, Flight.Distance, Age),
    ~ as.numeric(scale(.))
  ))
table_full_enc <- table_full_std %>%
  mutate(across(where(is.character), as.factor))
```

## Visualization

### Boxplots

Boxplots of each variable against satisfaction. As expected satisfied customers show greater averages in the service quality scores and lower averages for delays.

```{r fig.width=14, fig.height=25}

pairs <- lapply(all_vars, function(v) {
  c("Satisfaction", v)
})

plots <- lapply(pairs, function(p) {
  cat_var <- p[1]
  num_var <- p[2]

  ggplot(table_full_std, aes(x = .data[[cat_var]], y = .data[[num_var]])) +
  geom_boxplot() +
  labs(
    x = cat_var,
    y = num_var,
    title =""
  ) +
  theme_bw()
})

wrap_plots(plots, ncol = 3)
```

### Barplots

Just to show the number of observations of each value of the categorical variables

```{r fig.width=14, fig.height=25}

plot_bar <- function(vars, scale = NULL) {
  x_scale <- if (!is.null(scale)) {
    scale_x_discrete(limits = scale)
  } else {
    NULL
  }
  
  plots <- lapply(vars, function(v) {
    ggplot(table, aes(
      x = factor(.data[[v]]),
      fill = factor(Satisfaction)
    )) +
      
    geom_bar() +
      
    x_scale +
    scale_fill_manual(
      values = c(
        "Dissatisfied" = "red",
        "Satisfied" = "forestgreen"
      )
    ) +
      
    labs(
      title = v,
      x = NULL,
      y = "Count",
      fill = "Satisfaction"
    ) +
    theme_bw()
  })
}


plots <- plot_bar(c(target_var, cat_vars))
plots <- c(plots, plot_bar(score_vars, c("0", "1", "2", "3", "4", "5")))

wrap_plots(plots, ncol = 3)
```

### Barplots Percentages

Here we show the percentages of satisfied/dissatisfied customers for each value of each categorical variable. This is somewhat more informative than the normal barplots, tho the trends can be approximated visually in the graphs above.

Note: there are some odd values, e.g. 100% of the customers that gave Leg.Room = 0 are satisfied, but notice that there is only 1 such observation, thus it does not represent a statistically relevant phenomena -\> almost all scores show few observations for score = 0, so generally ignore the corresponding column in the following graphs.

So excluding the first column, we see that the percentage of customers that gave score = x increases as x increases, as expected.

The only two exceptions that still have a significant number of observations are `Food.and.Drink` and `Seat.Comfort`: those might be outliers, `score = 0` yet satisfied.

```{r fig.width=14, fig.height=20}

plot_bar_perc <- function(vars, scale = NULL) {
  x_scale <- if (!is.null(scale)) {
    scale_x_discrete(limits = scale)
  } else {
    NULL
  }
  
  plots <- lapply(vars, function(v) {

    ggplot(table, aes(
      x = factor(.data[[v]]),
      fill = Satisfaction
    )) +
    geom_bar(position = "fill") +
  
    geom_text(
      stat = "count",
      aes(label = after_stat(count)),
      position = position_fill(vjust = 0.5),
      color = "white",
      size = 3
    ) +
  
    scale_y_continuous(labels = scales::percent) +
    x_scale +
    scale_fill_manual(
      values = c(
        "Dissatisfied" = "red",
        "Satisfied" = "forestgreen"
      )
    ) +
      
    labs(
      title = v,
      x = NULL,
      y = "Percentage",
      fill = "Satisfaction"
    ) +
    theme_bw()
  })
}


plots <- c(plot_spacer(), plot_bar_perc(cat_vars))
plots <- c(plots, plot_bar_perc(score_vars, c("0", "1", "2", "3", "4", "5")))

wrap_plots(plots, ncol = 3)

```

### Pairplots

Using only the numerical variables, since the scores are just a whole lot of overlapping points.

Those are not very informative imo.

```{r fig.width=10, fig.height=10}

ggpairs(
  table_std,
  columns = num_vars,
  aes(color = Satisfaction, alpha = 0.5)
)
```

## Correlation

From the correlation matrix we see that most predictors are uncorrelated, with a few exceptions, such as

-   between the "Online" sth sth variables, which kinda makes sense, a good online service is good across all aspects.

-   between the flight enjoyment variables (Seat.Comfort, Food.and.Drink, Inflight.Entertainment)

-   `Departure.Delay` and `Arrival.Delay`: if you depart late, you will arrive late, thus almost full correlation.

and more with similar interpretations.

```{r fig.width=12, fig.height=12}

corr_matrix <- cor(table_full_std[, c(num_vars, score_vars)], use = "complete.obs")

corrplot(corr_matrix,
         method = "color",
         type = "upper",
         addCoef.col = "black",
         tl.col = "black",
         tl.srt = 45)
```

# Initial Regression Model

Start by fitting a regression model using only the predictors

$$\texttt{Age},\ \texttt{SeatComfort},\ \texttt{FlightDistance},\ \texttt{Class},\ \texttt{DepartureDelay},\ \texttt{ArrivalDelay}$$

First of all, BAS is using BMA (Bayesian Model Averaging), so we train many models and give them a weight.

To interpret the summary:

-   `P(B != 0 | Y)`: aka Posterior Inclusion Probability (PIP), how likely the variable belongs to the true model

-   The 1/0 under the models for the first rows (those corresponding to the predictors): boolean flag, whether that predictor is included in the model

-   `BF`: Bayes Factor, idk what it is but it's a metric that compares each model with the best one, the higher the better, the best model has `BF = 1` by construction.

-   `PostProb`: probability that that model is the true one, given the data

-   `logmar`: the higher the better

From the summary we see that:

-   `Flight.Distance` and `Departure.Delay` are not selected very frequently across all the models trained by BAS, they might not capture useful data to fit the `Satisfaction`

-   The best model obtained is much better than the others: `PostProbs` and `BF` are much higher than the other models, meaning that the data strongly prefers model 1 over the other alternatives.

-   `R2` is very bad for all the models, we are not fitting the data at all

In the summary of coefficients reports the average over the evaluated models weighted by their posterior probability, where

-   `post mean`: average effect of a variable across all models, but models where the variable is absent count as 0 effect, so this is a shrunken, uncertainty-aware estimate.

-   `post SD`: uncertainty on the effect, including the model uncertainty

so we can say that

-   `Seat.Comfort`: positive effect and high PIP, as expected

-   `ClassEco` and `ClassEco Plus`: they have negative effect (more likely to be dissatisfied when not in business class) and high PIP

-   `Arrival.Delay` and `Departure.Delay` have a negative effect (as expected) but a low PIP, however this is probably due to the veeery strong correlation between the two, so they split the PIP in two.

-   `Age`: poor PIP and a small coefficient, even if zero isn't in the 95% interval.

In the graph of the residuals the x is the prediction in the log-odds scale (I assume / ChatGPT says). I have to admit that i really can't understand that residual plot tho: when our model predicts log-odds = 0 (thus p(satisfied) = 0.5), we observe some residuals that are 0, which is impossible, any observations would either have satisfied = 0 or = 1, in both cases our prediction had a residual of 0.5. The same goes for residuals greater than 1. So maybe even the y is in the log-odds scale? but this still doesn't make sense, consider the right side, i guess (ChatGPT says) that the two lines correspond to the two values satisfaction can assume, then when we predict 2 (p = 0.88), we have a residual of -2 for one class and -1 for the other, this means that we should have predicted 4 or 3 (to be exact for one class) (or equivalently predict 0 and 1) , but this means making a prediction where we are more strongly convinced that satisfaction = 1 (or = 0 equivalently) for both classes =\> contradiction... i just don't know how to read it and ChatGPT isn't helping.

Note: all the values are generated with some randomness, so they can change slightly, but the interpretation should remain the same

```{r fig.width=10, fig.height=10}

fit <- bas.glm(
  Satisfaction ~ Age + Seat.Comfort + Flight.Distance + Class + Departure.Delay + Arrival.Delay,
  data = table_enc, family = binomial()
)

summary(fit)
coef(fit)
print("--------------------------------------------", quote = FALSE)
confint(coef(fit))

par(mfrow = c(2, 2))

plot(fit, which = 1)
plot(fit, which = 2)
plot(fit, which = 3)
plot(fit, which = 4)

par(mfrow = c(1, 1))

plot(confint(coef(fit)), main="coeff confidence intervals")
```

## Seat.Comfort as Categorical

Much slower but gives a better R2, however there are some weird things. We can expect a model deciding to exclude Seat.Comfort as a whole to make predictions, however here (since it's treated as 5 distinct variables) it can happen (as in model 1) that only a subset of the levels is selected, and others are ignored. Now, if the selected ones where all the k highest levels it would be fine (it means "if the score is above below a certain value, it might as well be treated as a 0 (which is the baseline level)"), however we see that level 5 is ignored, which means "if the observation has score = 5, it is equivalent to score = 0" which is clearly wrong. The solutions to this are three: group priors (idk what they are), hierarchical structure (idk what it is) or treating the factors as numeric, as done above.

This still has a better R2 (not that it was hard to be better than the previous model), but i still struggle to understand the residual plot.

```{r fig.width=8, fig.height=8}

table_cat <- table_enc %>%
  mutate(across(c(Seat.Comfort), as.factor))

fit_cat <- bas.glm(
  Satisfaction ~ Age + Seat.Comfort + Flight.Distance + Class + Departure.Delay + Arrival.Delay,
  data = table_cat, family = binomial()
)

summary(fit_cat)
coef(fit_cat)
print("--------------------------------------------", quote = FALSE)
confint(coef(fit_cat))

par(mfrow = c(2, 2))

plot(fit_cat, which = 1)
plot(fit_cat, which = 2)
plot(fit_cat, which = 3)
plot(fit_cat, which = 4)

par(mfrow = c(1, 1))

plot(confint(coef(fit_cat)),main="coeff confidence intervals")
```

# Full Model

The initial model used only 6 predictors. We now extend the analysis to all available covariates to identify which variables are most strongly associated with customer satisfaction.

With 21 predictors the full model space ($2^{21}$) is too large to enumerate exactly. `bas.glm` uses MCMC sampling over the model space, visiting models proportionally to their posterior probability.

```{r}
fit_full <- bas.glm(
  Satisfaction ~ . - Satisfaction - X,
  data            = table_enc,
  family          = binomial(),
  method          = "MCMC",
  MCMC.iterations = 10000
)

# ?? i don't know if we need to add:  prior = "BIC" 

round(summary(fit_full), 3)
```

**Image plot** : each column is one of the top-visited models (ranked by posterior probability); each row is a predictor. A coloured cell means the predictor is included in that model. Variables that appear coloured across almost all columns have high posterior inclusion probability and are consistently selected.

```{r fig.width=10, fig.height=8}
image(fit_full, rotate = FALSE)

```

**Posterior Inclusion Probabilities (PIPs)**: we extract and rank the PIPs to identify the most important predictors:

```{r}
pip_full <- data.frame(
  variable = fit_full$namesx,
  PIP      = fit_full$probne0,
  coeff    = coef(fit_full)$postmean
)
pip_full <- pip_full[order(pip_full$PIP, decreasing = TRUE), ]
pip_full
```

Variables with PIP close to 1 are consistently selected across nearly all visited models and should be considered the key drivers of satisfaction.

**Highest Posterior Model (HPM)**

```{r}
# Index of the model with larger log-likehood 
best_idx <- which.max(fit_full$logmarg)

# Considered variable
hpm_vars <- fit_full$which[[best_idx]]
cat("Variables in the model HPM (indexes):", hpm_vars, "\n")
cat("Name:", fit_full$namesx[hpm_vars + 1], "\n")

# Posterior probability of the model HPM vs runner-up
pp_sorted <- sort(fit_full$postprobs, decreasing = TRUE)
cat("\nPosterior probability HPM:      ", round(pp_sorted[1], 4), "\n")
cat("Posterior probability runner-up:", round(pp_sorted[2], 4), "\n")
```

```{r}
coef_full <- coef(fit_full, estimator = "BMA")
coef_full
print("--------------------------------------------", quote = FALSE)
confint(coef_full)
```

```{r fig.width=12, fig.height=10}
par(mfrow = c(4, 4))
plot(coef_full, ask = FALSE)
par(mfrow = c(1, 1))
```

```{r}
plot(confint(coef_full), main = "BMA — 95% credible intervals for coefficients")

```

```{r fig.width=10, fig.height=10}

par(mfrow = c(2, 2))

plot(fit_full, which = 1)
plot(fit_full, which = 2)
plot(fit_full, which = 3)
plot(fit_full, which = 4)

par(mfrow = c(1, 1))
```

**Interpretation of PIP and coefficients:**

BMA results, based on the top 2148 visited models, identify three groups of variables.

NOTE: the following depend heavily on the train-test split (in fact they are not up to date, just refer to the PIP table above), can we TODO sth about this variability? like stratification

**Strong drivers (PIP \> 0.95):** `Inflight.Entertainment` (+0.858), `Customer.TypeLoyal` (+2.306), `Ease.of.Online.Booking` (+0.455), `On.Board.Service` (+0.293), `Seat.Comfort` (+0.374) and `Checkin` (+0.214) all have positive effects, while `ClassEco` (−0.887), `Type.of.TravelPersonal` (−0.944) and `Departure.Arrival.Time.Convenient` (−0.222) are negative. All their credible intervals exclude or just touch zero. Notably, loyal customers show the largest effect by far (+2.306), and economy class passengers are the most penalised group (−0.887).

**Moderate evidence (0.5 \< PIP \< 0.95):** `Arrival.Delay` (−0.386), `Food.and.Drink` (−0.210), `Flight.Distance` (−0.190), `ClassEco Plus` (−0.677) and `Age` (−0.141) show some signal but with wider credible intervals that include zero, so their contribution is uncertain once the stronger predictors are in the model.

**Weak or redundant (PIP \< 0.5):** `Gate.Location`, `Inflight.WiFi`, `Leg.Room`, `Cleanliness`, `Online.Boarding`, `Baggage.Handling` and `Departure.Delay` are selected in fewer than half the visited models and their posterior means are close to zero. Notably, `Departure.Delay` (post mean = −0.002) is nearly entirely absorbed by `Arrival.Delay` once both are included.

# Variable Selection

In principle we could use the HPM as a form of variable selection, however it can be very unstable (as always due to the presence of correlated predictors), since it does not account for the model uncertainty, unless `PostProb` is concentrated in one model.

## PIP Threshold

Starting from the full model we see that many variables have low PIP, this confirms our belief that many predictors captured redundant information (e.g. `Arrival.Delay` is a consequence of `Departure.Delay`).

We can start by fitting a simpler model where we consider only the variables with PIP above a certain threshold.

Notice that

-   the PIP threshold might separate the dummy variables for categorical predictors =\> in the model I must use the original variable

-   now all predictors have high PIP, even those that had PIP slightly above the threshold in the full model, this is due to the removal of the correlated predictors that shared part of the PIP.

-   the model is now small enough to train it without MCMC

-   only one model (the one with all predictors) has a high `PostProb`, meaning that the predictions will be heavily influenced by that single model, but it's not that strong to state that there is no model uncertainty

```{r fig.width=8, fig.height=8}

# TODO: try other thresholds? 
#       maybe by looking at the sorted list of pip seeing if we missed sth important (that's not correlated to what we have selected)
#       or included sth useless

vars_th <- pip_full$variable[pip_full$PIP > 0.5]
vars_th <- setdiff(vars_th, "Intercept")
cat("Selected using PIP threshold: ", vars_th, "\n")

# TODO: automate the model construction, for the cat variables we should do sth like if (vars_th[i].starts_with(cat_vars[j for some j])) then use cat_vars[j]

fit_th <- bas.glm(
  Satisfaction ~
    Customer.Type + Inflight.Entertainment + Ease.of.Online.Booking + Class +
    Seat.Comfort + On.Board.Service + Type.of.Travel + Arrival.Delay + Food.and.Drink,
  data = table_enc, family = binomial()
)

summary(fit_th)
coef(fit_th)
print("--------------------------------------------", quote = FALSE)
confint(coef(fit_th))

par(mfrow = c(2, 2))

plot(fit_th, which = 1)
plot(fit_th, which = 2)
plot(fit_th, which = 3)
plot(fit_th, which = 4)

par(mfrow = c(1, 1))

plot(confint(coef(fit_th)), main="coeff confidence intervals")
```

This approach is simple, however we should compare carefully it's performance compared to the full model, because we have many strongly correlated predictors, so PIP might be split among them (e.g. one predictor could have PIP = 0.9, but the presence of a second predictor that captures the same information yields both PIP = 0.45). This means that we might have removed the predictors that captured that information, losing it.

For this reason we might want to manually choose a subset of the predictors excluding those that are strongly correlated to the selected ones and fit such a model.

## Manual Selection

The following variables have been chosen based on

-   the box-plots, picking the variables that show the most clear division between satisfied and dissatisfied customers,

-   the correlation matrix, trying not to pick two strongly correlated variables.

-   the knowledge of the problem: a loyal customer is more likely to be satisfied otherwise he wouldn't be loyal in the first place, delays sure lower the satisfaction and so on...

We see that overall the model is very simple and the best model obtained (that includes all the selected variables) has a much better `R2` compared to the suggested model. Being the `PostProb` all concentrated in one model means that BMA is kinda pointless in this case, all the other models obtained (each excluding a useful variable) have almost 0 weight. However this doesn't mean that we found the correct model, it just means that among all the models that contain the selected variables, we are sure that that's the "more adequate" one. So, when making predictions we can just use the HPM.

Now, this model isn't much simpler than the threshold one (just 1 less variable) (and not very different either) so we might leave this just as "sth we tested" or make it even simpler and see how it goes (but i wouldn't know what to remove, it already seems the bare minimum to make a decision about satisfaction).

```{r fig.width=8, fig.height=8}

fit_man <- bas.glm(
  Satisfaction ~
    Customer.Type + Arrival.Delay + Inflight.Entertainment + Online.Boarding +
    Cleanliness + Class + Type.of.Travel + Seat.Comfort,
  data = table_enc, family = binomial()
)

summary(fit_man)
coef(fit_man)
print("--------------------------------------------", quote = FALSE)
confint(coef(fit_man))

par(mfrow = c(2, 2))

plot(fit_man, which = 1)
plot(fit_man, which = 2)
plot(fit_man, which = 3)
plot(fit_man, which = 4)

par(mfrow = c(1, 1))

plot(confint(coef(fit_man)), main="coeff confidence intervals")
```

# Transforming Predictors

## Clustering Correlated Predictors

From the knowledge of the problem domain, we can also identify some clusters of features:

-   Digital experience cluster: `Inflight.WiFi`, `Inflight.Entertainment`, `Online.Support`, `Ease.of.Online.Booking`, `Online.Boarding`
-   On-board service cluster: `On.Board.Service`, `Leg.Room`, `Cleanliness`, `Baggage.Handling`, `Checkin`

Given the correlation matrix we see that the members of the first cluster are also highly correlated (maybe excluding `Inflight.Entertainment`) so we can try to merge them in a single feature that summarizes the whole effect. We could for instance take the average score across the predictors in the cluster (or better do PCA, but i refuse to think about ASS more than the strict necessary however).

?? maybe todo

## Other Transformations

?? Do we want to try sth for this, maybe squaring the scores can give a much stronger effect on the 5/5 scores, but do we suspect a quadratic relationship with the target? also the numeric features don't seem to have any particular pattern in the pairplots so i wouldn't transform them.

## Interaction Features

?? Do we suspect interactions? like idk `Flight.Entertainment` is relevant only when in business class or the combo `Seat.Comfort` and `Leg.Room` is stronger than the two individual effects so we should multiply them or sth like that... I think we have a looot of pussybilities (cit) so we should focus on the phenomena that we suspect are happening given our knowledge of planes. I know a lot about them from the outside, but ive never taken one so lemme know.

Motivation: Business class passengers have different expectations than Economy passengers. The effect of in-flight services on satisfaction may therefore depend on the travel class. We test two interactions: Class:Inflight.Entertainment and Class:Seat.Comfort

```{r}
fit_int <- bas.glm(
  Satisfaction ~
    Customer.Type + Inflight.Entertainment + Ease.of.Online.Booking + Class +
    Seat.Comfort + On.Board.Service + Type.of.Travel + Arrival.Delay +
    Class:Inflight.Entertainment + Class:Seat.Comfort,
  data            = table_enc,
  family          = binomial(),
  method          = "MCMC",
  MCMC.iterations = 10000
)

summary(fit_int)

coef(fit_int)
print("--------------------------------------------", quote = FALSE)
confint(coef(fit_int))

par(mfrow = c(2, 2))

plot(fit_int, which = 1)
plot(fit_int, which = 2)
plot(fit_int, which = 3)
plot(fit_int, which = 4)

par(mfrow = c(1, 1))

plot(confint(coef(fit_int)), main="coeff confidence intervals")
```

```{r}
pip_int <- data.frame(
  variable = fit_int$namesx,
  PIP      = fit_int$probne0,
  coeff    = coef(fit_int)$postmean
)
pip_int <- pip_int[order(pip_int$PIP, decreasing = TRUE), ]
pip_int
```

The PIP table reveals a clear separation between main effects and interaction terms. The main effects `Customer.TypeLoyal` (PIP = 0.999), `Inflight.Entertainment` (PIP = 1.000), `Ease.of.Online.Booking` (PIP = 0.999), `On.Board.Service` (PIP = 0.958) and `Arrival.Delay` (PIP = 0.995) are selected in virtually every visited model, confirming the results already obtained in the full model without interactions.

The interaction terms tell a more nuanced story. `Inflight.Entertainment:ClassEco` has a PIP of 0.656 and `Inflight.Entertainment:ClassEco Plus` of 0.527 — both above 0.5, suggesting moderate evidence that the effect of in-flight entertainment on satisfaction is not uniform across travel classes. Economy and Eco Plus passengers appear to respond differently to entertainment quality compared to Business class passengers, which is the reference category. By contrast, the `Seat.Comfort` interaction terms (`ClassEco:Seat.Comfort` PIP = 0.216, `ClassEco Plus:Seat.Comfort` PIP = 0.238) fall well below 0.5, indicating that the effect of seat comfort on satisfaction does not vary meaningfully across classes once the other predictors are accounted for.

The coefficient plot confirms the PIP results. `Customer.TypeLoyal` and `Inflight.Entertainment` show large positive effects with credible intervals well above zero. `ClassEco` has a clearly negative effect, consistent with the earlier finding that Economy passengers are less satisfied on average. `Arrival.Delay` retains a small but reliably negative effect. The interaction terms, not shown explicitly in the plot due to shrinkage from BMA, have posterior means close to zero, reflecting the moderate PIP values.

**Conclusion on interactions**

The evidence for interactions is moderate rather than conclusive. The `Inflight.Entertainment × Class` terms are selected in the majority of visited models, providing suggestive evidence that entertainment quality matters more — or less — depending on travel class. However, given the limited improvement in R² and predictive performance relative to the full main-effects model, the interaction model is best interpreted as an exploratory finding rather than a definitive improvement. The full model without interactions remains the preferred specification for prediction.

# Predictive Exercise

In this section we evaluate the predictive performance of the Bayesian logistic regression model. We split the dataset into a **learning set** (used to fit the model) and a **test set** (held out for prediction).

First define some useful functions since we have to evaluate multiple models

```{r}

# compute metrics such as accuracy
# - pred: output of "predict(model, ...)"
# - threshold: predictions higher than "threshold" are considered "Satisfied"
evaluate_performance <- function(pred, threshold, dataset) {
  prob <- pred$fit

  pred <- ifelse(prob >= threshold, "Satisfied", "Dissatisfied")
  pred <- factor(pred, levels = levels(dataset$Satisfaction))
  
  acc <- mean(pred == dataset$Satisfaction)
  conf <- table(Predicted = pred, Actual = dataset$Satisfaction)
  
  sens <- conf["Satisfied",    "Satisfied"]    / sum(conf[, "Satisfied"])
  spec <- conf["Dissatisfied", "Dissatisfied"] / sum(conf[, "Dissatisfied"])

  list(
    accuracy = acc,
    sensitivity = sens,
    specificity = spec,
    conf_mat = conf
  )
}

# plot the predicted probabilities of two models
# - pred_1, pred_2: output of two "predict(model, ...)" (they have to use the same prediction dataset)
compare_predictions <- function(pred_1, pred_2, dataset, labelx = "Predictions 1", labely = "Predictions 2") {
  prob_1 <- as.numeric(pred_1$fit)
  prob_2 <- as.numeric(pred_2$fit)
  
  pred_df <- data.frame(
    data_1 = prob_1,
    data_2 = prob_2,
    Actual = dataset$Satisfaction
  )
  
  ggplot(pred_df, aes(x = data_1, y = data_2, color = Actual)) +
    geom_point(alpha = 0.5, size = 1.5) +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
    scale_color_manual(
      values = c("Dissatisfied" = "red", "Satisfied" = "forestgreen")
    ) +
    xlab(labelx) +
    ylab(labely) +
    ggtitle("Predictions Comparison") +
    theme_bw()
}

# plot the predicted probabilities of BMA and HPM estimators of the same model
compare_bma_hpc <- function(model, dataset) {
  pred_bma <- predict(model, newdata = dataset, estimator = "BMA")
  pred_hpm <- predict(model, newdata = dataset, estimator = "HPM")
  compare_predictions(pred_hpm, pred_bma, dataset, "HPM predicted", "BMA predicted")
}

# Prints the perfromance metrics
# - performance: the output of "evaluate_performance(...)"
print_performance <- function(performance) {
  cat("Accuracy:   ", round(performance$accuracy, 4), "\n")
  cat("Sensitivity:", round(performance$sensitivity, 4), "(Satisfied correctly identified)\n")
  cat("Specificity:", round(performance$specificity, 4), "(Dissatisfied correctly identified)\n")
  
  cat("\nConfusion matrix:\n")
  print(performance$conf_mat)
}

# Prints the performance metrics for two models
# - perf_1, perf_2: the output of "evaluate_performance(...)"
print_performances <- function(perf_1, perf_2, name_1 = "Model 1", name_2 = "Model 2") {
  cat(sprintf("             %-6s %-6s\n", name_1, name_2))
  cat("Accuracy:   ", round(perf_1$accuracy, 4), round(perf_2$accuracy, 4), "\n")
  cat("Sensitivity:", round(perf_1$sensitivity, 4), round(perf_2$sensitivity, 4), "(Satisfied correctly identified)\n")
  cat("Specificity:", round(perf_1$specificity, 4), round(perf_2$specificity, 4), "(Dissatisfied correctly identified)\n")
  
  cat("\nConfusion Matrix ", name_1, ":\n", sep = "")
  print(perf_1$conf_mat)
  cat("\nConfusion Matrix ", name_2, ":\n", sep = "")
  print(perf_2$conf_mat)
}

compute_roc <- function(dataset, probs) {
  roc(
    response  = as.numeric(dataset$Satisfaction) - 1,
    predictor = as.numeric(probs)
  )
}

# Plots ROC curves
# - rocs: a "list(...)" of the values returned by "compute_roc(...)"
# - palette: a "c(...)" of colors at least as long as "rocs" or NULL
# - legend: a "c(...)" of strings to display in the legend
plot_roc <- function(rocs, palette = NULL, title = NULL, legend) {
  if (is.null(title))
    title = "ROC curves"
  if (is.null(palette))
    palette = c("steelblue", "darkorange", "forestgreen", "purple", "red", "yellow", "cyan", "blue")
  
  plot(rocs[[1]], col = palette[1], lwd = 2, xlim = c(1, 0), main = title,
       print.auc = TRUE, print.auc.y = 0.5, 
       print.thres = c(0.5), print.thres.adj = c(-0.05, 1.25), print.thres.pattern = "%.3f"
  )
  for (i in 2:length(rocs)) {
    plot(rocs[[i]], col = palette[i], lwd = 2, xlim = c(1, 0),
         print.auc = TRUE, print.auc.y = 0.5 - (i - 1) * 0.05,
         print.thres = c(0.5), print.thres.adj = c(-0.05, 1.25 + (i - 1) * 0.5), print.thres.pattern = "%.3f",
         add = TRUE
    )
  }
  legend("bottomright", legend = legend, col = palette, lwd = 2)
}

# Bar plot showing the selected metric for all trained models
# - metric: any scalar field in the return value of "evaluate_performance(...)"
plot_perf_comparison <- function(metric) {
  df <- data.frame(
    group = rep(c("Full", "Threshold", "Manual", "Initial"), each = 2),
    type  = rep(c("BMA", "HPM"), times = 4),
    value = c(test_full_bma_perf[[metric]], test_full_hpm_perf[[metric]],
              test_th_bma_perf[[metric]]  , test_th_hpm_perf[[metric]],
              test_man_bma_perf[[metric]] , test_man_hpm_perf[[metric]],
              test_init_bma_perf[[metric]], test_init_hpm_perf[[metric]])
  )
  
  ggplot(df, aes(x = group, y = value, fill = type)) +
    geom_col(position = "dodge") +
    coord_cartesian(ylim = c(0, 1)) +
    theme_minimal() +
    labs(x = "", y = "Value", fill = "Type")
}
```

## Full Model

### Predict Learning Set

We first obtain the fitted probabilities on the learning set (in-sample) and convert them to predicted classes using a 0.5 threshold, to get a baseline picture of how the model behaves on the data it was trained on.

```{r}
train_full_bma_pred <- predict(fit_full, estimator = "BMA")
train_full_hpm_pred <- predict(fit_full, estimator = "HPM")
train_full_bma_prob <- train_full_bma_pred$fit
```

```{r}
train_full_bma_perf <- evaluate_performance(train_full_bma_pred, 0.5, table_enc)
train_full_hpm_perf <- evaluate_performance(train_full_hpm_pred, 0.5, table_enc)

cat("Train set performance, BMA vs HPM:\n")
print_performances(train_full_bma_perf, train_full_hpm_perf, "BMA", "HPM")
```

**BMA vs HPM**

```{r}
compare_bma_hpc(fit_full, table_enc)
```

### Predict Test Set

We now predict on `test_enc`, i.e. the observations that were **not used to fit the model**. We obtain predictions under two estimators:

-   **BMA** (Bayesian Model Averaging): averages predictions across all visited models, weighted by their posterior probability;
-   **HPM** (Highest Posterior Model): uses only the single model with the largest posterior probability.

```{r}
test_full_bma_pred <- predict(fit_full, newdata = test_enc, estimator = "BMA")
test_full_hpm_pred <- predict(fit_full, newdata = test_enc, estimator = "HPM")

test_full_bma_prob <- test_full_bma_pred$fit
test_full_hpm_prob <- test_full_hpm_pred$fit
```

Below we plot the true response against the fitted and predicted probabilities. Black points are **in-sample** (learning set) and red points are **out-of-sample** (test set). A good model should show comparable spread between the two groups.

```{r}
# Combine in-sample and out-of-sample for a joint plot
plot(
  train_full_bma_prob,
  as.numeric(table_enc$Satisfaction) - 1.01,
  pch  = 16,
  xlab = expression(hat(p)),
  ylab = "Observed (0 = Dissatisfied, 1 = Satisfied)",
  main = "Fitted (black) vs predicted (red) probabilities"
)
points(
  test_full_bma_prob,
  as.numeric(test_enc$Satisfaction) - 0.99,
  pch = 16,
  col = "red"
)
abline(v = 0.5, lty = 2, col = "gray50")

```

-   The model is well-separated: most predictions are concentrated near$\hat{p}$​≈0 or $\hat{p}$​≈1, meaning the model is generally confident in its classifications rather than uncertain around 0.5. This is a sign that the selected predictors carry strong discriminative information.

-   The 0.5 threshold works well: points on y = 1 are predominantly on the right of the dashed line, and points on y = 0 are predominantly on the left, confirming that the 0.5 cutoff correctly separates the two classes in most cases.

-   Misclassified observations are those where y = 1 but $\hat{p}$​\<0.5 (satisfied customers predicted as dissatisfied, visible on the left side of the y = 1 line) and vice versa. These are relatively few, consistent with the accuracy we computed.

-   Red and black points overlap closely, indicating that the model generalises well from the learning set to the test set, with no obvious signs of overfitting.

**BMA vs HPM**

The scatter plot below compares the predicted probabilities under BMA and HPM on the test set. Points on the 45-degree dashed line indicate identical predictions; deviations show where the two estimators disagree. BMA predictions tend to be slightly more conservative (pulled toward 0.5) because they average over models rather than committing to a single one.

```{r}
compare_bma_hpc(fit_full, test_enc)
```

The plot compares the predicted probabilities of the two estimators on the test set, with points coloured by the true class.

-   **BMA and HPM agree closely**: almost all points lie very near the 45-degree dashed line, meaning the two estimators produce nearly identical predictions in this dataset. This is a reassuring sign: the data are informative enough that the posterior is concentrated around a single dominant model, so averaging across models (BMA) or picking the best one (HPM) leads to the same conclusion.

-   **BMA is slightly more conservative than HPM**: points tend to fall marginally *above* the diagonal for low predicted probabilities and *below* for high ones. In other words, BMA pulls predictions slightly toward 0.5 compared to HPM. This is the expected behaviour: BMA averages over models of varying complexity, and the simpler models included in the average shrink predictions toward the prior, while HPM commits fully to the single most probable model.

-   **The two classes are well separated**: green points (Satisfied) cluster in the upper-right region ( $\hat{p}$​ close to 1 for both estimators) and red points (Dissatisfied) cluster in the lower-left ( $\hat{p}$​ close to 0). The overlap between the two colours is limited, confirming the good discriminative ability of the model already observed in the previous plot.

-   **Most disagreement between BMA and HPM occurs in the intermediate range** ( $\hat{p}$​​≈0.25–0.60), which is exactly where model uncertainty matters most. Outside this range the predictions are nearly identical, because a predicted probability close to 0 or 1 is robust to which model is selected.

**Classification Performance**

We threshold the BMA and HPM probabilities at 0.5 and compare the resulting classifications with the true labels.

```{r}
test_full_bma_perf <- evaluate_performance(test_full_bma_pred, 0.5, test_enc)
test_full_hpm_perf <- evaluate_performance(test_full_hpm_pred, 0.5, test_enc)

cat("Test set performance, BMA vs HPM:\n")
print_performances(test_full_bma_perf, test_full_hpm_perf, "BMA", "HPM")
```

The out-of-sample results confirm that both estimators generalise well to unseen data, with BMA outperforming HPM slightly.

**BMA achieves an accuracy of 85.50%**, correctly classifying 342 out of 400 test observations. **HPM achieves 84.00%**, correctly classifying 336 out of 400. The gap is modest but consistent with the theoretical expectation: by averaging over models rather than committing to one, BMA absorbs model uncertainty and produces more robust predictions.

Looking at the confusion matrices in detail:

-   Under **BMA**, 152 dissatisfied customers are correctly identified and 190 satisfied ones are correctly identified. The errors are 26 false positives (dissatisfied predicted as satisfied) and 32 false negatives (satisfied predicted as dissatisfied), roughly balanced between the two types of error.

-   Under **HPM**, the picture is very similar: 148 and 188 correct classifications respectively, with 30 false positives and 34 false negatives. HPM makes slightly more false positive errors than BMA (30 vs 26), which explains most of the accuracy gap.

-   The **sensitivity of BMA is 0.8559**, meaning that 85.59% of truly satisfied customers are correctly identified. The **specificity is 0.8539**, meaning that 85.39% of truly dissatisfied customers are correctly identified. The two metrics are very close to each other, which indicates that the model is **well-balanced**: it does not systematically favour one class over the other, a desirable property when both types of misclassification carry comparable costs.

Overall, an accuracy above 80% with balanced sensitivity and specificity on held-out data, using a model estimated purely on Bayesian principles with no manual tuning of the decision threshold, is a strong result and confirms that the selected predictors carry genuine and replicable signal about passenger satisfaction.

**AUC**

The **Area Under the ROC Curve (AUC)** evaluates the model's discriminative ability across *all possible classification thresholds*, not just the 0.5 cutoff used for accuracy. The ROC curve plots sensitivity (true positive rate) against 1 − specificity (false positive rate) as the threshold varies from 0 to 1. An AUC of 0.5 corresponds to a random classifier, while an AUC of 1 indicates perfect separation between the two classes.

```{r}
# AUC — BMA
test_full_bma_roc <- compute_roc(test_enc, test_full_bma_prob)
test_full_bma_auc <- auc(test_full_bma_roc)

# AUC — HPM
test_full_hpm_roc <- compute_roc(test_enc, test_full_hpm_prob)
test_full_hpm_auc <- auc(test_full_hpm_roc)

cat("BMA - AUC:", round(test_full_bma_auc, 4), "\n")
cat("HPM - AUC:", round(test_full_hpm_auc, 4), "\n")
```

Both estimators achieve excellent discriminative performance. BMA reaches an AUC of **0.9134** and HPM of **0.9059**, both well above 0.9, a threshold commonly used to indicate *outstanding* classification ability.

```{r}
plot_roc(rocs = list(test_full_bma_roc, test_full_hpm_roc), legend = c("BMA", "HPM"), title = "Full - Test")
```

The ROC curves confirm this: they rise steeply toward the upper-left corner of the plot, meaning the model achieves high sensitivity while keeping the false positive rate low, across virtually every threshold. This is consistent with the well-separated predicted probability distributions observed earlier.

The gap between BMA and HPM is small (0.0075), but BMA consistently outperforms HPM both in AUC and in accuracy (85.50% vs 84.00%). This is the expected theoretical behaviour: by averaging predictions across all visited models weighted by their posterior probability, BMA absorbs model uncertainty and produces more robust, better-calibrated probabilities than committing to a single model as HPM does.

## Initial Model

### Predict Learning Set

```{r}
train_init_bma_pred <- predict(fit, estimator = "BMA")
train_init_hpm_pred <- predict(fit, estimator = "HPM")
train_init_bma_prob <- train_full_bma_pred$fit
```

```{r}
train_init_bma_perf <- evaluate_performance(train_init_bma_pred, 0.5, table_enc)
train_init_hpm_perf <- evaluate_performance(train_init_hpm_pred, 0.5, table_enc)

cat("Train set performance, BMA vs HPM:\n")
print_performances(train_init_bma_perf, train_init_hpm_perf, "BMA", "HPM")
```

**BMA vs HPM**

```{r}
compare_bma_hpc(fit, table_enc)
```

### Predict Test Set

```{r}
test_init_bma_pred <- predict(fit, newdata = test_enc, estimator = "BMA")
test_init_hpm_pred <- predict(fit, newdata = test_enc, estimator = "HPM")

test_init_bma_prob <- test_init_bma_pred$fit
test_init_hpm_prob <- test_init_hpm_pred$fit
```

```{r}
# Combine in-sample and out-of-sample for a joint plot
plot(
  train_init_bma_prob,
  as.numeric(table_enc$Satisfaction) - 1.01,
  pch  = 16,
  xlab = expression(hat(p)),
  ylab = "Observed (0 = Dissatisfied, 1 = Satisfied)",
  main = "Fitted (black) vs predicted (red) probabilities"
)
points(
  test_init_bma_prob,
  as.numeric(test_enc$Satisfaction) - 0.99,
  pch = 16,
  col = "red"
)
abline(v = 0.5, lty = 2, col = "gray50")

```

**BMA vs HPM**

```{r}
compare_bma_hpc(fit, test_enc)
```

**Classification Performance**

```{r}
test_init_bma_perf <- evaluate_performance(test_init_bma_pred, 0.5, test_enc)
test_init_hpm_perf <- evaluate_performance(test_init_hpm_pred, 0.5, test_enc)

cat("Test set performance, BMA vs HPM:\n")
print_performances(test_init_bma_perf, test_init_hpm_perf, "BMA", "HPM")
```

**AUC**

```{r}
test_init_bma_roc <- compute_roc(test_enc, test_init_bma_prob)
test_init_hpm_roc <- compute_roc(test_enc, test_init_hpm_prob)

plot_roc(rocs = list(test_init_bma_roc, test_init_hpm_roc), legend = c("BMA", "HPM"), title = "Initial - Test")
```

## Threshold Selection Model

### Predict Learning Set

```{r}
train_th_bma_pred <- predict(fit_th, estimator = "BMA")
train_th_hpm_pred <- predict(fit_th, estimator = "HPM")
train_th_bma_prob <- train_full_bma_pred$fit
```

```{r}
train_th_bma_perf <- evaluate_performance(train_th_bma_pred, 0.5, table_enc)
train_th_hpm_perf <- evaluate_performance(train_th_hpm_pred, 0.5, table_enc)

cat("Train set performance, BMA vs HPM:\n")
print_performances(train_th_bma_perf, train_th_hpm_perf, "BMA", "HPM")
```

**BMA vs HPM**

```{r}
compare_bma_hpc(fit_th, table_enc)
```

### Predict Test Set

```{r}
test_th_bma_pred <- predict(fit_th, newdata = test_enc, estimator = "BMA")
test_th_hpm_pred <- predict(fit_th, newdata = test_enc, estimator = "HPM")

test_th_bma_prob <- test_th_bma_pred$fit
test_th_hpm_prob <- test_th_hpm_pred$fit
```

```{r}
# Combine in-sample and out-of-sample for a joint plot
plot(
  train_th_bma_prob,
  as.numeric(table_enc$Satisfaction) - 1.01,
  pch  = 16,
  xlab = expression(hat(p)),
  ylab = "Observed (0 = Dissatisfied, 1 = Satisfied)",
  main = "Fitted (black) vs predicted (red) probabilities"
)
points(
  test_th_bma_prob,
  as.numeric(test_enc$Satisfaction) - 0.99,
  pch = 16,
  col = "red"
)
abline(v = 0.5, lty = 2, col = "gray50")

```

**BMA vs HPM**

```{r}
compare_bma_hpc(fit_th, test_enc)
```

**Classification Performance**

```{r}
test_th_bma_perf <- evaluate_performance(test_th_bma_pred, 0.5, test_enc)
test_th_hpm_perf <- evaluate_performance(test_th_hpm_pred, 0.5, test_enc)

cat("Test set performance, BMA vs HPM:\n")
print_performances(test_th_bma_perf, test_th_hpm_perf, "BMA", "HPM")
```

**AUC**

```{r}
test_th_bma_roc <- compute_roc(test_enc, test_th_bma_prob)
test_th_hpm_roc <- compute_roc(test_enc, test_th_hpm_prob)

plot_roc(rocs = list(test_th_bma_roc, test_th_hpm_roc), legend = c("BMA", "HPM"), title = "Threshold - Test")
```

## Manual Selection Model

### Predict Learning Set

```{r}
train_man_bma_pred <- predict(fit_man, estimator = "BMA")
train_man_hpm_pred <- predict(fit_man, estimator = "HPM")
train_man_bma_prob <- train_full_bma_pred$fit
```

```{r}
train_man_bma_perf <- evaluate_performance(train_man_bma_pred, 0.5, table_enc)
train_man_hpm_perf <- evaluate_performance(train_man_hpm_pred, 0.5, table_enc)

cat("Train set performance, BMA vs HPM:\n")
print_performances(train_man_bma_perf, train_man_hpm_perf, "BMA", "HPM")
```

**BMA vs HPM**

```{r}
compare_bma_hpc(fit_man, table_enc)
```

### Predict Test Set

```{r}
test_man_bma_pred <- predict(fit_man, newdata = test_enc, estimator = "BMA")
test_man_hpm_pred <- predict(fit_man, newdata = test_enc, estimator = "HPM")

test_man_bma_prob <- test_man_bma_pred$fit
test_man_hpm_prob <- test_man_hpm_pred$fit
```

```{r}
# Combine in-sample and out-of-sample for a joint plot
plot(
  train_man_bma_prob,
  as.numeric(table_enc$Satisfaction) - 1.01,
  pch  = 16,
  xlab = expression(hat(p)),
  ylab = "Observed (0 = Dissatisfied, 1 = Satisfied)",
  main = "Fitted (black) vs predicted (red) probabilities"
)
points(
  test_man_bma_prob,
  as.numeric(test_enc$Satisfaction) - 0.99,
  pch = 16,
  col = "red"
)
abline(v = 0.5, lty = 2, col = "gray50")

```

**BMA vs HPM**

```{r}
compare_bma_hpc(fit_man, test_enc)
```

**Classification Performance**

```{r}
test_man_bma_perf <- evaluate_performance(test_man_bma_pred, 0.5, test_enc)
test_man_hpm_perf <- evaluate_performance(test_man_hpm_pred, 0.5, test_enc)

cat("Test set performance, BMA vs HPM:\n")
print_performances(test_man_bma_perf, test_man_hpm_perf, "BMA", "HPM")
```

**AUC**

```{r}
test_man_bma_roc <- compute_roc(test_enc, test_man_bma_prob)
test_man_hpm_roc <- compute_roc(test_enc, test_man_hpm_prob)

plot_roc(rocs = list(test_man_bma_roc, test_man_hpm_roc), legend = c("BMA", "HPM"), title = "Manual - Test")
```

## Comparisons

Predicted probabilities

```{r fig.width=10, fig.height=8}

plots <- compare_predictions(test_full_bma_pred, test_th_bma_pred, test_enc, "Full - BMA - Test", "Threshold - BMA - Test")
plots <- c(plots, compare_predictions(test_full_bma_pred, test_man_bma_pred, test_enc, "Full - BMA - Test", "Manual - BMA - Test"))
plots <- c(plots, compare_predictions(test_full_bma_pred, test_init_bma_pred, test_enc, "Full - BMA - Test", "Initial - BMA - Test"))
plots <- c(plots, compare_predictions(test_th_bma_pred, test_man_bma_pred, test_enc, "Threshold - BMA - Test", "Manual - BMA - Test"))

wrap_plots(plots, ncol = 2)
```

ROC

```{r}
plot_roc(
  rocs = list(test_full_bma_roc, test_th_bma_roc, test_man_bma_roc, test_init_bma_roc), 
  legend = c("Full", "Threshold", "Manual", "Initial"), 
  title = "BMA - Test"
)
```

Accuracy and friends

```{r}
plot_perf_comparison("accuracy")
plot_perf_comparison("sensitivity")
plot_perf_comparison("specificity")
```

# Comparison Beta-Prior

- **CCH (default)** — `CCH(a = 0.5, b = 1, s = 0)`-\>places a diffuse prior on the shrinkage factor `g/(1+g)`, adapting moderately to the data without strongly favouring either sparse or complex models. A good general-purpose baseline when no prior belief about model size is available

- **AIC** — `aic.prior()` -\> fixes `g = exp(2) ≈ 7.39`, making the prior on coefficients nearly uninformative so that the likelihood dominates. Posterior model probabilities behave similarly to AIC weights, resulting in light penalisation of complexity and a tendency to include more variables. The least selective of the four priors

- **BIC** — `bic.prior(n = nrow(table_enc))` -\> sets `g = n`, placing strong prior mass near zero and penalising model complexity proportionally to `log(n)`, in analogy with the BIC criterion. More selective than `aic.prior()`: variables need a stronger association with the response to achieve high PIP. Consistent as `n → ∞`, but can be overly conservative in finite samples

- **Robust** — `robust()` -\> rather than fixing `g`, places a mixture prior over it that adapts to the fit of each model. Avoids the information paradox of fixed g-priors, is consistent, and performs reliably across a wide range of sample sizes and signal strengths. The recommended default when no strong external information on model complexity is available

```{r}
# Initial model 
prior_configs <- list(
  list(name = "CCH (default)", betaprior = NULL),
  list(name = "AIC",    betaprior = aic.prior()),
  list(name = "BIC",    betaprior = bic.prior(n = nrow(table_enc))),
  list(name = "Robust", betaprior = robust())
)

formula_init <- Satisfaction ~ Age + Seat.Comfort + Flight.Distance + Class + Departure.Delay + Arrival.Delay

# Fit all the models and save the results
prior_fits <- lapply(prior_configs, function(cfg) {
  if (is.null(cfg$betaprior)) {
    bas.glm(formula_init, data = table_enc, family = binomial())
  } else {
    bas.glm(formula_init, data = table_enc, family = binomial(), betaprior = cfg$betaprior)
  }
})
names(prior_fits) <- sapply(prior_configs, `[[`, "name")
```

## Pip Comparison

```{r}
# PIP TABLE — compare Posterior Inclusion Probabilities

pip_list <- lapply(names(prior_fits), function(nm) {
  fit_i <- prior_fits[[nm]]
  data.frame(
    variable = fit_i$namesx,
    PIP      = round(fit_i$probne0, 3),
    prior    = nm
  )
})

pip_compare <- do.call(rbind, pip_list)

# one clumn per prior
pip_wide <- reshape(
  pip_compare,
  idvar     = "variable",
  timevar   = "prior",
  direction = "wide"
)
names(pip_wide) <- gsub("PIP\\.", "", names(pip_wide))
rownames(pip_wide) <- NULL
pip_wide <- pip_wide[order(pip_wide[["CCH (default)"]], decreasing = TRUE), ]

cat("=== Posterior Inclusion Probabilities per prior ===\n")
print(pip_wide, row.names = FALSE)
```
```{r}
# PLOT — PIP per viariable

# Esclude intercept
pip_plot_df <- pip_compare[pip_compare$variable != "Intercept", ]

ggplot(pip_plot_df, aes(x = reorder(variable, PIP), y = PIP, fill = prior)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "gray40") +
  coord_flip() +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title = "Posterior Inclusion Probabilities — comparison betaprior",
    x     = NULL,
    y     = "PIP",
    fill  = "Prior"
  ) +
  theme_bw()
```
## BMA coefficient

```{r}
# BMA COEFFICIENT — posterior mean e SD for all the prior

coef_list <- lapply(names(prior_fits), function(nm) {
  cf    <- coef(prior_fits[[nm]])
  ci    <- confint(cf)
  data.frame(
    variable  = cf$namesx,
    post.mean = round(cf$postmean,  4),
    post.sd   = round(cf$postsd,    4),
    lower.95  = round(ci[, 1],      4),
    upper.95  = round(ci[, 2],      4),
    prior     = nm
  )
})
coef_compare <- do.call(rbind, coef_list)

for (v in unique(coef_compare$variable)) {
  sub <- coef_compare[coef_compare$variable == v, c("prior", "post.mean", "post.sd", "lower.95", "upper.95")]
  cat("\n--- Variable:", v, "---\n")
  print(sub, row.names = FALSE)
}
```
```{r}
# PLOT — Credible intervals comparison across priors
coef_plot_df <- coef_compare[coef_compare$variable != "Intercept", ]

ggplot(coef_plot_df, aes(x = prior, y = post.mean, color = prior)) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = lower.95, ymax = upper.95), width = 0.3) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray40") +
  facet_wrap(~ variable, scales = "free_y", ncol = 3) +
  scale_color_brewer(palette = "Set2") +
  labs(
    title = "Posterior means and 95% credible intervals by prior",
    x     = NULL,
    y     = "Posterior mean",
    color = "Prior"
  ) +
  theme_bw() +
  theme(
    axis.text.x  = element_blank(),
    axis.ticks.x = element_blank()
  )
```
**Posterior Coefficients — Comparison Across Priors**

The posterior means and standard deviations of the coefficients are broadly stable across the four priors, which is reassuring: the main conclusions about the direction and magnitude of each effect do not depend on the prior choice.

**Class** (both `ClassEco` and `ClassEco Plus`) shows the strongest and most consistent effect across all priors, with posterior means around −1.66 and −1.63 respectively and small standard deviations. Economy and Eco Plus passengers are substantially less likely to be satisfied than Business class passengers regardless of the prior used, and this finding is robust.

**Seat.Comfort** is the only positive predictor in this model, with a posterior mean of approximately 0.45 across all four priors and negligible variation. Higher seat comfort scores are consistently associated with greater satisfaction.

**Age** is the most prior-sensitive variable: the posterior mean ranges from 0.048 under BIC to 0.117 under AIC, reflecting the fact that its association with satisfaction is weak and the prior shrinkage materially affects the estimate. BIC, being the most selective prior, shrinks it the most toward zero.

**Flight.Distance**, **Departure.Delay** and **Arrival.Delay** all show negative posterior means, as expected, but with relatively large standard deviations. The two delay variables in particular show high variability across priors — most notably AIC produces larger magnitudes and wider uncertainty for both — which reflects their strong mutual correlation: the signal is shared between them and neither is selected consistently on its own.

Overall, the four priors lead to essentially the same substantive interpretation. The only meaningful differences appear for the weakest predictors (Age, the two delays), where the degree of shrinkage depends on how conservatively the prior treats model complexity.

## Performance

```{r}
# PERFORMANCE — Accuracy, Sensitivity, Specificity, AUC

perf_summary <- do.call(rbind, lapply(names(prior_fits), function(nm) {
  fit_i <- prior_fits[[nm]]
  
  # Prediction BMA on test set
  pred_test <- predict(fit_i, newdata = test_enc, estimator = "BMA")
  prob_test  <- pred_test$fit
  
  perf <- evaluate_performance(pred_test, 0.5, test_enc)
  
  # AUC
  roc_i <- compute_roc(test_enc, prob_test)
  
  data.frame(
    Prior       = nm,
    Accuracy    = round(perf$accuracy,    4),
    Sensitivity = round(perf$sensitivity, 4),
    Specificity = round(perf$specificity, 4),
    AUC         = round(as.numeric(auc(roc_i)), 4)
  )
}))

cat("=== Performance on Test Set (BMA, threshold 0.5) ===\n")
print(perf_summary, row.names = FALSE)
```
**Predictive Performance — Comparison Across Priors**

Overall, the four priors yield very similar predictive performance on the test set, with accuracy ranging narrowly between 0.74 and 0.757 and AUC between 0.781 and 0.784. This confirms that, for this model and dataset, the choice of prior has little practical impact on out-of-sample prediction.

**BIC achieves the highest accuracy (0.757) and sensitivity (0.789)**, meaning it correctly classifies more observations overall and identifies the largest share of truly satisfied customers. However, it does so at the cost of the lowest specificity (0.716), so it is slightly more prone to misclassifying dissatisfied customers as satisfied. This trade-off is consistent with the behaviour of a more selective prior: by including fewer variables and shrinking coefficients more aggressively, BIC produces predicted probabilities that are somewhat more concentrated, shifting the effective decision boundary.

**CCH, AIC and Robust perform almost identically**, with accuracy around 0.740 and nearly equal sensitivity and specificity. Among these three, CCH achieves marginally better accuracy (0.743) while AIC and Robust reach the highest AUC (0.784 and 0.783 respectively), suggesting slightly better rank discrimination even if the threshold-based classification is equivalent.

**The AUC values are all close to 0.78**, indicating moderate but not outstanding discriminative ability. This is expected given that the model uses only the six suggested covariates — the full model trained on all available predictors achieves an AUC above 0.91, confirming that the initial six-variable model leaves substantial predictive signal unexploited.

In summary, no prior dominates clearly across all metrics. BIC is preferable if sensitivity is prioritised, while CCH or Robust offer a more balanced trade-off between sensitivity and specificity and are theoretically better motivated as general-purpose choices.

## ROC

```{r}
# PLOT — ROC curves

roc_list <- lapply(names(prior_fits), function(nm) {
  pred_i <- predict(prior_fits[[nm]], newdata = test_enc, estimator = "BMA")
  compute_roc(test_enc, pred_i$fit)
})

plot_roc(
  rocs   = roc_list,
  legend = names(prior_fits),
  title  = "ROC — Comparison betaprior"
)
```
**ROC Curves — Comparison Across Priors**

The ROC curves for all four priors overlap almost perfectly throughout the entire specificity-sensitivity range, forming a single indistinguishable band. This is the visual confirmation of what the performance table already suggested: the choice of prior has essentially no impact on the discriminative ability of the model, with all four AUC values clustering tightly around 0.78.

## Comparison BMA with default

```{r}
pred_default <- predict(prior_fits[["CCH (default)"]], newdata = test_enc, estimator = "BMA")$fit

plots_prior <- lapply(setdiff(names(prior_fits), "CCH (default)"), function(nm) {
  pred_i <- predict(prior_fits[[nm]], newdata = test_enc, estimator = "BMA")$fit
  
  df <- data.frame(
    CCH     = as.numeric(pred_default),
    other   = as.numeric(pred_i),
    Actual  = test_enc$Satisfaction
  )
  
  ggplot(df, aes(x = CCH, y = other, color = Actual)) +
    geom_point(alpha = 0.5, size = 1.5) +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
    scale_color_manual(values = c("Dissatisfied" = "red", "Satisfied" = "forestgreen")) +
    labs(
      title = paste("CCH vs", nm),
      x     = "CCH (default)",
      y     = nm
    ) +
    theme_bw()
})

wrap_plots(plots_prior, ncol = 2)
```
**Predicted Probabilities — CCH vs Other Priors**

In all three scatter plots the points lie almost exactly on the 45-degree diagonal, indicating that CCH, AIC, BIC and Robust produce virtually identical predicted probabilities for every observation in the test set. There is no visible systematic deviation for either class: satisfied (green) and dissatisfied (red) points track the diagonal equally well across the full probability range.

This confirms that, for this six-variable model, the prior specification has no meaningful effect on the predicted probabilities, and all four choices can be considered interchangeable from a predictive standpoint

# Comparison g prior

- **g = 1** : the prior variance equals the information in a single observation. Coefficients are pulled very strongly toward zero, and the prior dominates unless the likelihood provides overwhelming evidence. This leads to very low PIP for most variables and heavily shrunk posterior means. In practice it is rarely a sensible choice unless the analyst has strong prior belief that all effects are negligible

- **g = 10 :** a moderate value that balances prior and likelihood. Shrinkage is still present but less aggressive than g=1, and variables with a genuine but modest association with the response can achieve reasonable PIP. ROughly comparable in behaviour to aic.prior()

- **g = 100 :** the prior is now quite diffuse, assigning non-negligible probability to large coefficient values. The likelihood dominates and most variables tend to be included. PIP values are generally higher than for smaller g, and posterior means are closer to the MLE. The risk is mild overfitting of the model space

- **g = n :** setting `g` equal to the sample size is known as the **unit-information prior**: the prior carries the same amount of information as a single observation from the dataset. This is exactly the choice underlying `bic.prior()`, and posterior model probabilities approximate BIC weights

```{r}
# COMPARISON G-PRIOR — g = 1, 10, n, 100

g_configs <- list(
  list(name = "g = 1",   betaprior = g.prior(g = 1)),
  list(name = "g = 10",  betaprior = g.prior(g = 10)),
  list(name = "g = n",   betaprior = g.prior(g = nrow(table_enc))),
  list(name = "g = 100", betaprior = g.prior(g = 100))
)

g_fits <- lapply(g_configs, function(cfg) {
  bas.glm(formula_init, data = table_enc, family = binomial(), betaprior = cfg$betaprior)
})
names(g_fits) <- sapply(g_configs, `[[`, "name")
```

## Pip comparison

```{r}
# PIP TABLE

pip_g_list <- lapply(names(g_fits), function(nm) {
  fit_i <- g_fits[[nm]]
  data.frame(
    variable = fit_i$namesx,
    PIP      = round(fit_i$probne0, 3),
    prior    = nm
  )
})

pip_g_compare <- do.call(rbind, pip_g_list)

pip_g_wide <- reshape(
  pip_g_compare,
  idvar     = "variable",
  timevar   = "prior",
  direction = "wide"
)
names(pip_g_wide) <- gsub("PIP\\.", "", names(pip_g_wide))
rownames(pip_g_wide) <- NULL
pip_g_wide <- pip_g_wide[order(pip_g_wide[["g = 1"]], decreasing = TRUE), ]

cat("=== PIP — g-prior comparison ===\n")
print(pip_g_wide, row.names = FALSE)
```
```{r}
# PLOT — PIP barplot confronto g

pip_g_plot_df <- pip_g_compare[pip_g_compare$variable != "Intercept", ]
pip_g_plot_df$prior <- factor(pip_g_plot_df$prior,
                              levels = c("g = 1", "g = 10", "g = 100", "g = n"))

ggplot(pip_g_plot_df, aes(x = reorder(variable, PIP), y = PIP, fill = prior)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "gray40") +
  coord_flip() +
  scale_fill_brewer(palette = "Set1") +
  labs(
    title = "Posterior Inclusion Probabilities — g-prior comparison",
    x     = NULL,
    y     = "PIP",
    fill  = "g"
  ) +
  theme_bw()
```
**PIP — G-Prior Comparison**

The PIP table reveals a clear pattern: as `g` increases, posterior inclusion probabilities generally decrease for the weaker predictors, while the strongest ones remain robust across all values of `g`.

**Seat.Comfort**, **ClassEco** and **ClassEco Plus** maintain PIP close to 1.000 across all four values of `g`, confirming that these are the dominant drivers of satisfaction in this model. Their inclusion is essentially certain regardless of the prior specification.

**Flight.Distance**, **Age**, **Departure.Delay** and **Arrival.Delay** show much greater sensitivity to `g`. Under `g = 1`, all four have PIP above 0.85, but this is an artefact of the very strong shrinkage: when all coefficients are forced close to zero, even weak predictors appear relevant to absorb residual variation. As `g` increases to 10 and then to `n`, the PIP of these variables drops substantially — most notably `Age` falls from 0.889 to 0.323 and `Arrival.Delay` from 0.849 to 0.375 — revealing that their association with satisfaction is genuinely weak once the prior allows stronger effects for the truly important variables.

The behaviour at `g = 100` is somewhat irregular: PIP values are higher than at `g = n` for several variables, which is unexpected since larger `g` should in principle be less selective. This is likely due to the stochastic nature of the BAS enumeration rather than a genuine prior effect.

Overall, the results confirm that **Seat.Comfort** and **Class** are the only reliable predictors in this six-variable model, while the remaining four covariates carry limited and prior-sensitive signal.

## BMA coefficient

```{r}
# COEFFICIENTS — posterior mean, SD, 95% credible intervals

coef_g_list <- lapply(names(g_fits), function(nm) {
  cf <- coef(g_fits[[nm]])
  ci <- confint(cf)
  data.frame(
    variable  = cf$namesx,
    post.mean = round(cf$postmean, 4),
    post.sd   = round(cf$postsd,   4),
    lower.95  = round(ci[, 1],     4),
    upper.95  = round(ci[, 2],     4),
    prior     = nm
  )
})

coef_g_compare <- do.call(rbind, coef_g_list)

for (v in unique(coef_g_compare$variable)) {
  sub <- coef_g_compare[coef_g_compare$variable == v,
                        c("prior", "post.mean", "post.sd", "lower.95", "upper.95")]
  cat("\n--- Variable:", v, "---\n")
  print(sub, row.names = FALSE)
}
```

```{r}
# PLOT — Credible intervals

coef_g_plot_df <- coef_g_compare[coef_g_compare$variable != "Intercept", ]

coef_g_plot_df$prior <- factor(coef_g_plot_df$prior,
                               levels = c("g = 1", "g = 10", "g = 100", "g = n"))

ggplot(coef_g_plot_df, aes(x = prior, y = post.mean, color = prior)) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = lower.95, ymax = upper.95), width = 0.3) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray40") +
  facet_wrap(~ variable, scales = "free_y", ncol = 3) +
  scale_color_brewer(palette = "Set1") +
  labs(
    title = "Posterior means and 95% credible intervals — g-prior comparison",
    x     = NULL,
    y     = "Posterior mean",
    color = "g"
  ) +
  theme_bw() +
  theme(
    axis.text.x  = element_blank(),
    axis.ticks.x = element_blank()
  )
```
**Posterior Coefficients — G-Prior Comparison**

The results show a consistent pattern across variables: as `g` increases from 1 to larger values, posterior means move away from zero and credible intervals widen, reflecting the progressively weaker shrinkage imposed by the prior.

**Seat.Comfort** shows the most striking shrinkage effect: under `g = 1` the posterior mean is only 0.225, roughly half the value obtained under `g = 10`, `g = n` and `g = 100` (all around 0.44–0.45). This is a direct consequence of the very strong shrinkage at `g = 1`, which pulls even well-supported coefficients toward zero. Under all other values of `g` the estimates are consistent and the credible intervals exclude zero, confirming a robust positive effect.

**ClassEco** and **ClassEco Plus** display the same pattern: the magnitude of the negative effect is substantially underestimated at `g = 1` (around −0.83 for both) compared to larger values of `g` (around −1.62 to −1.67), where the credible intervals are tighter and entirely below zero. This confirms that `g = 1` distorts the estimates of even the strongest predictors.

**Age**, **Flight.Distance**, **Departure.Delay** and **Arrival.Delay** all have credible intervals that include zero under most values of `g`, consistent with their low PIP. The upper bounds for `Age` are positive but close to zero, and the delay variables show wide and symmetric intervals, reflecting high uncertainty about their individual contributions given their mutual correlation.

Overall, `g = 1` produces systematically shrunk and unreliable estimates and should be avoided. The estimates under `g = 10`, `g = n` and `g = 100` are broadly consistent with each other, with `g = n` being the theoretically preferred choice among fixed-g options.

## Performance

```{r}
# PERFORMANCE — Accuracy, Sensitivity, Specificity, AUC

perf_g_summary <- do.call(rbind, lapply(names(g_fits), function(nm) {
  fit_i      <- g_fits[[nm]]
  pred_test  <- predict(fit_i, newdata = test_enc, estimator = "BMA")
  prob_test  <- pred_test$fit
  perf       <- evaluate_performance(pred_test, 0.5, test_enc)
  roc_i      <- compute_roc(test_enc, prob_test)
  
  data.frame(
    Prior       = nm,
    Accuracy    = round(perf$accuracy,    4),
    Sensitivity = round(perf$sensitivity, 4),
    Specificity = round(perf$specificity, 4),
    AUC         = round(as.numeric(auc(roc_i)), 4)
  )
}))

cat("=== Performance on Test Set — g-prior comparison ===\n")
print(perf_g_summary, row.names = FALSE)
```
**Performance on Test Set — G-Prior Comparison**

Overall, the four g-prior specifications yield similar predictive performance, with accuracy ranging from 0.733 to 0.757 and AUC values tightly clustered between 0.781 and 0.784. As with the betaprior comparison, the choice of g has limited practical impact on out-of-sample prediction.

**g = n achieves the highest accuracy (0.757) and sensitivity (0.789)**, confirming it as the best-performing fixed-g choice, consistent with its theoretical motivation as the unit-information prior. Its specificity (0.716) is identical to `g = 10`, meaning the gain in accuracy comes from better identification of satisfied customers without sacrificing the ability to detect dissatisfied ones.

**g = 1 shows the most unbalanced behaviour**: it achieves the highest sensitivity (0.813) but the lowest specificity (0.634), meaning it tends to classify most observations as satisfied regardless of the true label. This is a direct consequence of the strong shrinkage at `g = 1`, which flattens predicted probabilities and shifts the effective decision boundary, producing a model biased toward the majority class.

**g = 10 and g = 100** perform similarly to each other and to the betaprior results, with accuracy around 0.740 and a more balanced trade-off between sensitivity and specificity.

**AUC values are nearly identical across all four settings**, ranging from 0.781 to 0.784, confirming that rank discrimination is unaffected by the choice of g. The performance differences observed in accuracy and sensitivity are therefore purely a threshold effect rather than a genuine difference in discriminative ability.

## ROC

```{r}
# PLOT — ROC curves

roc_g_list <- lapply(names(g_fits), function(nm) {
  pred_i <- predict(g_fits[[nm]], newdata = test_enc, estimator = "BMA")
  compute_roc(test_enc, pred_i$fit)
})

plot_roc(
  rocs   = roc_g_list,
  legend = names(g_fits),
  title  = "ROC — g-prior comparison"
)
```
**ROC Curves — G-Prior Comparison**

The four ROC curves overlap almost perfectly across the entire specificity-sensitivity range, with all AUC values equal to 0.783. This is the visual confirmation of what the performance table already showed: the choice of `g` has no meaningful impact on the discriminative ability of the model.

## Comparison BMA with different g priors

```{r}
pred_gn <- predict(g_fits[["g = n"]], newdata = test_enc, estimator = "BMA")$fit

plots_g <- lapply(setdiff(names(g_fits), "g = n"), function(nm) {
  pred_i <- predict(g_fits[[nm]], newdata = test_enc, estimator = "BMA")$fit
  
  df <- data.frame(
    g_n    = as.numeric(pred_gn),
    other  = as.numeric(pred_i),
    Actual = test_enc$Satisfaction
  )
  
  ggplot(df, aes(x = g_n, y = other, color = Actual)) +
    geom_point(alpha = 0.5, size = 1.5) +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
    scale_color_manual(values = c("Dissatisfied" = "red", "Satisfied" = "forestgreen")) +
    labs(
      title = paste("g = n vs", nm),
      x     = "g = n",
      y     = nm
    ) +
    theme_bw()
})

wrap_plots(plots_g, ncol = 2)
```
**Predicted Probabilities — G-Prior Comparison**

**G = n vs g = 10** and **g = n vs g = 100** show points lying very close to the 45-degree diagonal for both classes, indicating that these three specifications produce nearly identical predicted probabilities across all test observations. The agreement is essentially the same as observed in the betaprior comparison, confirming that once `g` is large enough to let the data dominate, the exact value has little effect on predictions.

**G = n vs g = 1** tells a markedly different story. The points deviate substantially from the diagonal, with `g = 1` predictions compressed into a much narrower range (roughly 0.35 to 0.80) compared to `g = n` (0.10 to 0.95). This is the direct consequence of the strong shrinkage at `g = 1`: predicted probabilities are pulled toward 0.5, making the model less confident in its classifications. The two classes are still separated but with considerably less margin, which explains the lower specificity observed in the performance table.

Overall, the plots confirm that `g = 1` is the only specification that produces meaningfully different predictions, and in a detrimental way. The choices `g = 10`, `g = n` and `g = 100` are essentially interchangeable from a predictive standpoint, with `g = n` remaining the preferred choice on theoretical grounds

# Comparison Model prior

- **uniform** : Assigns equal probability to all 2p2\^p 2p possible models. It makes no assumption about model size, treating a model with 1 variable exactly as likely a priori as a model with all 6. It is the natural reference point but can implicitly favour medium-sized models since there are combinatorially more of them

- **Beta- Binomial:** places a Beta-Binomial prior on the number of variables included, introducing uncertainty about model size. It is more flexible than Uniform because it adapts to the data also in terms of complexity, and it is the most commonly recommended choice in the BAS literature for general use

- **Truncated Beta-Binomial:** a truncated version of the Beta-Binomial that excludes the null model (no predictors) and the saturated model (all predictors) from the model space. Useful when it is certain that at least one variable should be included and that not all variables are relevant simultaneously

- **Truncated Poisson:** places a truncated Poisson prior on the number of included variables, favouring sparse models more aggressively than the Beta-Binomial. Particularly suited when strong prior belief exists that only a small number of predictors are truly relevant.

```{r}
#rm(beta.binomial)
p <- 6 #number of coviarates
n <- nrow(table_enc)

model_prior_configs <- list(
  list(name = "Uniform",       modelprior = BAS::uniform()),
  list(name = "Beta-Binomial", modelprior = BAS::beta.binomial(alpha = 1, beta = 1)),
  list(name = "Tr-Beta-Binom", modelprior = BAS::tr.beta.binomial(alpha = 1, beta = 1, trunc = p)),
  list(name = "Tr-Poisson",    modelprior = BAS::tr.poisson(lambda = 5, trunc = p))
)
model_prior_fits <- lapply(model_prior_configs, function(cfg) {
  bas.glm(
    formula_init,
    data       = table_enc,
    family     = binomial(),
    modelprior = cfg$modelprior
  )
})
names(model_prior_fits) <- sapply(model_prior_configs, `[[`, "name")
```

## Pip comparison

```{r}
# PIP TABLE
 
pip_mp_list <- lapply(names(model_prior_fits), function(nm) {
  fit_i <- model_prior_fits[[nm]]
  data.frame(
    variable = fit_i$namesx,
    PIP      = round(fit_i$probne0, 3),
    prior    = nm
  )
})

pip_mp_compare <- do.call(rbind, pip_mp_list)

pip_mp_wide <- reshape(
  pip_mp_compare,
  idvar     = "variable",
  timevar   = "prior",
  direction = "wide"
)
names(pip_mp_wide) <- gsub("PIP\\.", "", names(pip_mp_wide))
rownames(pip_mp_wide) <- NULL
pip_mp_wide <- pip_mp_wide[order(pip_mp_wide[["Uniform"]], decreasing = TRUE), ]

cat("=== PIP — model prior comparison ===\n")
print(pip_mp_wide, row.names = FALSE)
```
```{r}
# COEFFICIENTS — posterior mean, SD, 95% credible intervals

coef_mp_list <- lapply(names(model_prior_fits), function(nm) {
  cf <- coef(model_prior_fits[[nm]])
  ci <- confint(cf)
  data.frame(
    variable  = cf$namesx,
    post.mean = round(cf$postmean, 4),
    post.sd   = round(cf$postsd,   4),
    lower.95  = round(ci[, 1],     4),
    upper.95  = round(ci[, 2],     4),
    prior     = nm
  )
})

coef_mp_compare <- do.call(rbind, coef_mp_list)

for (v in unique(coef_mp_compare$variable)) {
  sub <- coef_mp_compare[coef_mp_compare$variable == v,
                         c("prior", "post.mean", "post.sd", "lower.95", "upper.95")]
  cat("\n--- Variable:", v, "---\n")
  print(sub, row.names = FALSE)
}
```
```{r}
# PLOT — Credible intervals per variabile, confronto model prior

coef_mp_plot_df <- coef_mp_compare[coef_mp_compare$variable != "Intercept", ]

ggplot(coef_mp_plot_df, aes(x = prior, y = post.mean, color = prior)) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = lower.95, ymax = upper.95), width = 0.3) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray40") +
  facet_wrap(~ variable, scales = "free_y", ncol = 3) +
  scale_color_brewer(palette = "Set1") +
  labs(
    title = "Posterior means and 95% credible intervals — model prior comparison",
    x     = NULL,
    y     = "Posterior mean",
    color = "Model Prior"
  ) +
  theme_bw() +
  theme(
    axis.text.x  = element_blank(),
    axis.ticks.x = element_blank()
  )
```
```{r}
# PLOT — PIP barplot confronto model prior

pip_mp_plot_df <- pip_mp_compare[pip_mp_compare$variable != "Intercept", ]

ggplot(pip_mp_plot_df, aes(x = reorder(variable, PIP), y = PIP, fill = prior)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "gray40") +
  coord_flip() +
  scale_fill_brewer(palette = "Set1") +
  labs(
    title = "Posterior Inclusion Probabilities — model prior comparison",
    x     = NULL,
    y     = "PIP",
    fill  = "Model Prior"
  ) +
  theme_bw()
```
## Performance

```{r}
# PERFORMANCE — Accuracy, Sensitivity, Specificity, AUC

perf_mp_summary <- do.call(rbind, lapply(names(model_prior_fits), function(nm) {
  fit_i     <- model_prior_fits[[nm]]
  pred_test <- predict(fit_i, newdata = test_enc, estimator = "BMA")
  prob_test <- pred_test$fit
  perf      <- evaluate_performance(pred_test, 0.5, test_enc)
  roc_i     <- compute_roc(test_enc, prob_test)
  
  data.frame(
    Prior       = nm,
    Accuracy    = round(perf$accuracy,    4),
    Sensitivity = round(perf$sensitivity, 4),
    Specificity = round(perf$specificity, 4),
    AUC         = round(as.numeric(auc(roc_i)), 4)
  )
}))

cat("=== Performance on Test Set — model prior comparison ===\n")
print(perf_mp_summary, row.names = FALSE)
```
## ROC

```{r}
# PLOT — ROC curves

roc_mp_list <- lapply(names(model_prior_fits), function(nm) {
  pred_i <- predict(model_prior_fits[[nm]], newdata = test_enc, estimator = "BMA")
  compute_roc(test_enc, pred_i$fit)
})

plot_roc(
  rocs   = roc_mp_list,
  legend = names(model_prior_fits),
  title  = "ROC — model prior comparison"
)
```
```{r}
# PLOT — Predicted probabilities vs Uniform (default)

pred_uniform <- predict(model_prior_fits[["Uniform"]], 
                        newdata   = test_enc, 
                        estimator = "BMA")$fit

plots_mp <- lapply(setdiff(names(model_prior_fits), "Uniform"), function(nm) {
  pred_i <- predict(model_prior_fits[[nm]], newdata = test_enc, estimator = "BMA")$fit
  
  df <- data.frame(
    Uniform = as.numeric(pred_uniform),
    other   = as.numeric(pred_i),
    Actual  = test_enc$Satisfaction
  )
  
  ggplot(df, aes(x = Uniform, y = other, color = Actual)) +
    geom_point(alpha = 0.5, size = 1.5) +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
    scale_color_manual(values = c("Dissatisfied" = "red", "Satisfied" = "forestgreen")) +
    labs(
      title = paste("Uniform vs", nm),
      x     = "Uniform",
      y     = nm
    ) +
    theme_bw()
})

wrap_plots(plots_mp, ncol = 2)
```

